### 2023-08-11 draft_v0.Rmd
Initial commit. Downloaded data. Data prepartation complete and missingness explored. From Linux.

---
title: "Untitled"
output: html_document
date: "2023-08-11"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height = 16, fig.width = 20)
```

## Overview

```{r setup2}

# libraries
library(tidyverse)

# colours and theme

# functions

```

```{r data_prep}

# read file from local

dta <- read_csv("data/sr_hex_truncated.csv")

glimpse(dta)

# prepare ----------------------------------------------------------
  # make month and year columns

dta <- dta %>% 
  mutate(year_creation = lubridate::year(creation_timestamp),
         month_creation = lubridate::month(creation_timestamp),
         year_completion = lubridate::year(completion_timestamp),
         month_completion = lubridate::month(completion_timestamp))

mnths <- c("jan", "feb", "mar", "apr", "may", "jun", "jul",
           "aug", "sep", "oct", "nov", "dec")

unique(dta$month_creation)
dta$month_creation <- factor(dta$month_creation, levels = 1:12,  labels = mnths)
dta$month_completion <- factor(dta$month_completion, levels = 1:12, labels = mnths)

```

```{r data_checks}
sum(is.na(dta$notification_number))
# no missing notification nrs

length(dta$notification_number) == length(unique(dta$notification_number))
# each row is unique entry

# missingness

x <- nrow(dta)

map_dbl(dta, ~sum(is.na(.x))/x) %>% 
  sort()

dta %>% 
  group_by(is.na(directorate), is.na(department), is.na(branch), 
           is.na(section), is.na(official_suburb)) %>% 
  count()

unique(dta$section)
glimpse(dta)
```

### Description of the data and preparation steps

#### Missingness

We investigated missingness that would make interpretation hard.  Entries for which any of the Directorate, Department, Branch, Section or Official suburb were missing are visualised (Figure A). Entries with missing Directorate were in general also missing for the other categories. Entries for which Directorate was missing, but Official suburb was not(n = 1256), accounted for at most 4% of entries for any suburb and the suburb with the highest percentage of missing entries of this kind, was Kewtown(Figure B).

```{r missing1}
tmp <- dta %>% 
  select(directorate, department, branch, section, official_suburb) %>% 
  mutate(across(everything(), ~as.numeric(is.na(.x)))) %>% 
  as.data.frame()

na_interpret_affected <- UpSetR::upset(tmp)

na_interpret_affected
grid::grid.text("Fig A",x = 0.65, y=0.95, gp=grid::gpar(fontsize=20))

plotdta <- dta %>% 
  group_by(official_suburb) %>% 
  count(name = "total_suburb") %>% 
  inner_join(dta %>% 
  filter(is.na(directorate) & !is.na(official_suburb)) %>% 
  group_by(official_suburb) %>% 
  count(name = "n_missing")) %>% 
  mutate(percentage_missing = n_missing/total_suburb*100) %>% 
  arrange(desc(n_missing), desc(percentage_missing)) %>% 
  ungroup()

plotdta %>% 
  slice(1:20) %>% 
  ggplot()+
  geom_point(aes(percentage_missing, reorder(official_suburb, percentage_missing), size = n_missing))+
  labs(title = "Fig. B: Top 20 Suburbs with missing Directorate")

```

Of the entries with Official suburb missing, ...

```{r}

plotdta <- dta %>% 
  group_by(directorate) %>% 
  count(name = "total_directorate") %>%
  inner_join(dta %>% 
               filter(is.na(official_suburb)) %>% 
               group_by(directorate) %>% 
                        count(name = "n_missing")) %>% 
  ungroup() %>% 
  mutate(percentage_missing = n_missing/total_directorate*100)

plotdta %>% 
  ggplot()+
  geom_point(aes(percentage_missing, reorder(directorate, percentage_missing), size = n_missing))+
  labs(title = "Fig. C: Missing Suburbs per Directorate")

```

```{r overview}

dta %>% 
  select(directorate) %>% 
  arrange(directorate) %>% 
  distinct()

# missing directorate

dta %>% 
  filter(!complete.cases(dta)) %>% 
  group_by(is.na(.)) %>% 
  count() %>% 
  print(n = Inf)

dta %>% 
  filter(is.na(directorate)) %>% 
  select(-contains(c("number", "stamp", "year", "month")), -longitude, -latitude) %>% 
  distinct() %>% 
  print(width = Inf)
  group_by(official_suburb) %>% 
  count() %>% 
  arrange(desc(n))

```

## "In which 3 suburbs should the Urban Mobility directorate concentrate their infrastructure improvement efforts?"

## 1. Focusing on the Urban Mobility directorate - "What is the median & 80th percentile time to complete each service request across the City?" (each row represent a service request).

## 2. Focusing on the Urban Mobility directorate - "What is the median & 80th percentile time to complete each service request for the 3 suburbs identified in (1)?" (each row represent a service request).

## 3. "Is there any significant differences in the median and 80th percentile completion times between the City as a whole and the 3 suburbs identified in(1)?".  Please elaborate on the similarities or differences.

## 3. Provide a visual mock of a dashboard for the purpose of monitoring progress in applying the insights developed in (1) & (2). It should focus the user on performance pain points. Add a note for each visual element, explaining how it helps fulfill this overall function. Please also provide a brief explanation as to how the data provided would be used to realise what is contained in your mock.

## 4. Identify value-adding insights for the management of Urban Mobility, from the dataset provided, in regard to commuter transport within the City.

### 2023-08-11 draft_v0.Rmd
Exploration started. From Windows.

@@
 # libraries
+library(kableExtra)
+library(knitr)
 library(tidyverse)
@@
 
-unique(dta$section)
-glimpse(dta)
+# dates
+
 ```
@@
 
+# missing directorate and dates
+
+unique(dta$month_creation)
+
+dta %>% 
+  group_by(year_creation, month_creation) %>% 
+  count(name = "total_ym") %>% 
+  inner_join(dta %>% 
+               group_by(year_creation, month_creation, is.na(directorate)) %>% 
+               count(name = "n_missing")) %>%
+  ungroup() %>% 
+  rename(missing_directorate = 4) %>% 
+  filter(missing_directorate == TRUE) %>% 
+  mutate(percentage_missing = n_missing/total_ym*100) %>% 
+  kable() %>%
+  kable_styling(full_width = FALSE, position = "left")
+
 ```
@@
 
+# Exclude if missing directorate
+
+dta <- dta %>% 
+  filter(!is.na(directorate))
+
+# Missingness in urban mobility
+unique(dta$directorate)
+
+dta_um <- dta %>% 
+  filter(directorate == "URBAN MOBILITY")
+
+x <- which(names(dta_um) %in% c("code_group", "code", "cause_code_group", "cause_code"))
+
+map(dta_um[x], ~sort(unique(.x))) 
+map(dta_um[x], ~sum(is.na(.x)))
+
+dta_um %>%
+  select(all_of(x)) %>% 
+  group_by(is.na(.)) %>% 
+  count()
+
+x <- nrow(dta_um)
+
+x1 <- dta_um %>% 
+  group_by(is.na(cause_code_group), is.na(cause_code)) %>% 
+  count(name = "missing_cause") %>% 
+  ungroup() %>% 
+  rename(...1 = 1, ...2 = 2) %>% 
+  filter(...1 == TRUE & ...2 == TRUE) %>% 
+  pull(missing_cause)
+
+tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of entries.")) %>% 
+  kable(col.names = NULL) %>%
+  kable_styling(full_width = FALSE, position = "left")
+
 ```
@@
 ```{r overview}
+x <- nrow(dta)
 
 dta %>% 
-  select(directorate) %>% 
-  arrange(directorate) %>% 
-  distinct()
-
-# missing directorate
+  group_by(directorate) %>% 
+  count(name = "Nr_of_entries") %>% 
+  mutate(Percentage_of_total = round(Nr_of_entries/x*100, 2)) %>% 
+  arrange(desc(Nr_of_entries)) %>% 
+  kable(title = "Number and percentage of entries per Directorate") %>%
+  kable_styling(full_width = FALSE, position = "left")
 
 dta %>% 
-  filter(!complete.cases(dta)) %>% 
-  group_by(is.na(.)) %>% 
-  count() %>% 
-  print(n = Inf)
-
-dta %>% 
-  filter(is.na(directorate)) %>% 
-  select(-contains(c("number", "stamp", "year", "month")), -longitude, -latitude) %>% 
-  distinct() %>% 
-  print(width = Inf)
-  group_by(official_suburb) %>% 
-  count() %>% 
-  arrange(desc(n))
+  group_by(year_creation, month_creation) %>% 
+  count(name = "total_ym") %>% 
+  inner_join(dta %>% 
+               group_by(year_creation, month_creation, directorate) %>% 
+               count(name = "Nr_of_entries")) %>% 
+  ungroup() %>% 
+  mutate(Percentage_of_total = round(Nr_of_entries/total_ym*100, 2)) %>% 
+  arrange(desc(Nr_of_entries)) %>% 
+  ggplot(aes(month_creation, Percentage_of_total, 
+             fill = reorder(directorate, Percentage_of_total)))+
+  geom_col()+
+  facet_wrap(~year_creation, scales = "free_x") # resize left bar?
 

### 2023-08-11 draft_v0.Rmd
Explored completion times.  From Windows.

@@
 na_interpret_affected
-grid::grid.text("Fig A",x = 0.65, y=0.95, gp=grid::gpar(fontsize=20))
+grid::grid.text("Fig. A",x = 0.65, y=0.95, gp=grid::gpar(fontsize=20))
 
@@
 
-tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of entries.")) %>% 
+tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of Urban Mobility entries.")) %>% 
   kable(col.names = NULL) %>%
@@
   arrange(desc(Nr_of_entries)) %>% 
+  rename(Directorate = directorate) %>% 
   kable(title = "Number and percentage of entries per Directorate") %>%
@@
 
+```{r}
+
+```
+
 ## 1. Focusing on the Urban Mobility directorate - "What is the median & 80th percentile time to complete each service request across the City?" (each row represent a service request).
+
+```{r}
+
+# completion time missing
+x <- nrow(dta_um)
+
+dta_um %>% 
+  group_by(is.na(completion_timestamp)) %>% 
+  count(name = "Nr_missing") %>% 
+  ungroup() %>% 
+  rename(Missing_completion_time = 1) %>% 
+  mutate(across(Missing_completion_time, ~ifelse(.x == TRUE, "yes", "no")),
+         Percentage = round(Nr_missing/x*100, 2)) %>% 
+  kable() %>%
+  kable_styling(full_width = FALSE, position = "left")
+
+# breakdown of entries with missing completion times
+
+x <- c("department", "branch", "section", "code_group",
+  "code", "cause_code_group", "cause_code", "official_suburb")
+
+tmp <- map(x, 
+    ~dta_um %>% 
+  filter(is.na(completion_timestamp)) %>% 
+  select(notification_number, all_of(.x)) %>% 
+  mutate(across(-notification_number, ~paste(cur_column(), .x, sep = ":")),
+         value = 1) %>% 
+    rename(key = 2) %>% 
+  pivot_wider(names_from = key, values_from = value)) %>% 
+  purrr::reduce(inner_join) %>% 
+  as.data.frame()
+
+tmp[is.na(tmp)] <- 0
+
+UpSetR::upset(tmp)
+
+# exclude data with no completion time
+
+dta_um <- dta_um %>% 
+  filter(!is.na(completion_timestamp))
+
+```
+
+```{r}
+
+plotdta <- dta_um %>% 
+  mutate(diff_time_hrs = as.integer(completion_timestamp - creation_timestamp)/60) %>% 
+  select(notification_number, diff_time_hrs, code_group, code, cause_code_group,
+         cause_code, official_suburb)
+
+plot_time_per_var <- function(dataset = plotdta, column = "code_group", hours_cutoff = 96){
+
+  x <- which(names(dataset) == column)
+  dataset <- dataset %>% 
+    rename(col_select = x)
+
+  tmp <- dataset %>% 
+  group_by(col_select) %>% 
+  summarise(n = n(),
+            med_hrs = median(diff_time_hrs),
+            q5 = quantile(diff_time_hrs, probs = 0.05),
+            q80 = quantile(diff_time_hrs, probs = 0.8),
+            q95 = quantile(diff_time_hrs, probs = 0.95)) %>% 
+  ungroup()
+
+p1 <- tmp %>% 
+  filter(q95 < hours_cutoff) %>% 
+  select(col_select) %>% 
+  inner_join(dataset %>% 
+               select(col_select, diff_time_hrs)) %>% 
+  ggplot(aes(diff_time_hrs, col_select))+
+  ggdist::stat_interval()+
+  labs(y = column)
+
+p2 <- tmp %>% 
+  filter(q5 >= hours_cutoff) %>% 
+  select(col_select) %>% 
+  inner_join(dataset) %>% 
+  mutate(time_weeks = (diff_time_hrs/24)/7) %>% 
+  ggplot(aes(time_weeks, col_select))+
+  ggdist::stat_interval()+
+  labs(y = column)
+
+return(list(hours = p1, weeks = p2))
+
+}
+
+plotlist <- plot_time_per_var(dataset = plotdta, column = "code_group", hours_cutoff = 96)
+
+ggpubr::ggarrange(plotlist = plotlist)
+
+plotlist <- plot_time_per_var(dataset = plotdta, column = "code", hours_cutoff = 96)
+
+ggpubr::ggarrange(plotlist = plotlist)
+
+plotlist <- plot_time_per_var(dataset = plotdta, column = "cause_code_group", hours_cutoff = 96)
+
+ggpubr::ggarrange(plotlist = plotlist)
+
+plotlist <- plot_time_per_var(dataset = plotdta, column = "cause_code", hours_cutoff = 96)
+
+ggpubr::ggarrange(plotlist = plotlist)
+
+plotdta %>% 
+  ggplot(aes(diff_time_hrs, official_suburb))+
+  ggdist::stat_interval()
+
+ggpubr::ggarrange(plotlist = plotlist)
+
+```
 

### 2023-08-11 draft_v0.Rmd
Explore top 3. From Windows.

@@
 ```{r}
+# department, branch, section, code_group not helpful
+dta_um %>% 
+  group_by(code, cause_code) %>% 
+  count() %>% 
+  print(n = Inf)
+
+# filter out cause codes "Not Defined" etc, but keep NA
+
+dta_um <- dta_um %>% 
+  filter(!cause_code %in% c("No Cause Found",
+                           "None/Private") &
+           !is.na(official_suburb))
+
+dta_um %>% 
+  group_by(official_suburb) %>% 
+  count() %>% 
+  arrange(desc(n)) %>% 
+  print(n = 20)
+
+# top 3 by month
+
+plotdta <- dta_um %>% 
+  group_by(year_creation, month_creation, official_suburb) %>% 
+  count() %>% 
+  arrange(year_creation, month_creation, desc(n)) %>% 
+  print(n = 20) %>% 
+  ungroup()
+
+plotdta <- map(unique(plotdta$month_creation),
+           ~plotdta %>% 
+  filter(month_creation == .x) %>% 
+  arrange(desc(n)) %>% 
+  slice(1:10)) %>% 
+    bind_rows()
+
+dta_um %>% 
+  group_by(official_suburb) %>% 
+  count(name = "total_overall") %>% 
+  inner_join(plotdta) %>% 
+  ggplot(aes(n, official_suburb, fill = total_overall))+
+  geom_col()+
+  facet_wrap(~month_creation)
+
+# START HERE
 

### 2023-08-12 draft_v0.Rmd
More exploring of times. From Windows.

@@
 
-# top 3 by month
+# top 10 by month
 
@@
 
-# START HERE
+# looks like parklands, milnerton, table view, marconi beam
+
+# if look at median across months
+dta_um %>% 
+  group_by(month_creation, official_suburb) %>% 
+  count() %>% 
+  ungroup() %>% 
+  group_by(official_suburb) %>% 
+  summarise(med = median(n)) %>% 
+  ungroup() %>% 
+  arrange(desc(med))
+
+x <- c("PARKLANDS", "MARCONI BEAM", "BRACKENFELL COMMON",
+       "TABLE VIEW", "MILNERTON")
+
+dta_um %>% 
+  filter(official_suburb %in% x) %>% 
+  group_by(official_suburb, code) %>% 
+  count() %>% 
+  ungroup() %>% 
+  ggplot(aes(n, code, fill = official_suburb))+
+  geom_col(position = "dodge")
 
@@
 ```{r}
+# get back no fault found etc.
+
+dta_um <- dta %>% 
+  filter(directorate == "URBAN MOBILITY")
 
@@
   group_by(is.na(completion_timestamp)) %>% 
-  count(name = "Nr_missing") %>% 
+  count(name = "Nr") %>% 
   ungroup() %>% 
@@
   mutate(across(Missing_completion_time, ~ifelse(.x == TRUE, "yes", "no")),
-         Percentage = round(Nr_missing/x*100, 2)) %>% 
+         Percentage = round(Nr/x*100, 2)) %>% 
   kable() %>%
@@
 
-# exclude data with no completion time
+# exclude data with no completion time, create diff var
 
 dta_um <- dta_um %>% 
-  filter(!is.na(completion_timestamp))
+  filter(!is.na(completion_timestamp)) %>% 
+   mutate(time_hrs = as.integer(completion_timestamp - creation_timestamp)/60)
+
+```
+
+```{r}
+#  create table withs medians, q80 for all suburbs and for all of city
+
+time_tables_make <- function(dataset = dta_um, time_cutoff = NULL){
+
+  if(is.null(time_cutoff)){
+
+  tabledta <- dataset %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+  mutate(official_suburb = "ALL COMBINED") %>% 
+  bind_rows(dataset %>% 
+  group_by(official_suburb) %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+    ungroup()) %>% 
+  mutate(code = "all_codes_combined") %>% 
+  bind_rows(dataset %>% 
+  group_by(official_suburb, code) %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+    ungroup())
+
+  return(tabledta)
+
+} else if(!is.null(time_cutoff)){
+
+  dataset_x <- dataset %>% 
+    filter(time_hrs < time_cutoff)
+
+  tabledta_below <- dataset_x %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+  mutate(official_suburb = "ALL COMBINED") %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb) %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+    ungroup()) %>% 
+  mutate(code = "all_codes_combined") %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb, code) %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+    ungroup())
+
+  dataset_x <- dataset %>% 
+    filter(time_hrs >= time_cutoff)
+
+  tabledta_above <- dataset_x %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+  mutate(official_suburb = "ALL COMBINED") %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb) %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+    ungroup()) %>% 
+  mutate(code = "all_codes_combined") %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb, code) %>% 
+  summarise(med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+    ungroup())
+
+  return(list(tabledta_below = tabledta_below,
+                tabledta_above = tabledta_above))
+}
+}
+
+time_dta_list <- list()
+
+# for all of city, all codes, all completion time lengths
+
+time_dta_list$all_times <- time_tables_make(dataset = dta_um, time_cutoff = NULL)
+
+# for all of city, all codes, split by times < or > 96hrs
+tmp <- time_tables_make(dataset = dta_um, time_cutoff = 96)
+
+time_dta_list$below_96 <- tmp$tabledta_below 
+time_dta_list$above_eq_96 <- tmp$tabledta_above
+
+# weekends, public holidays?
 
@@
 
+# START HERE
 plot_time_per_var <- function(dataset = plotdta, column = "code_group", hours_cutoff = 96){
@@
   tmp <- dataset %>% 
+  group_by(col_select) %>% 
+  summarise(n = n(),
+            med_hrs = median(diff_time_hrs),
+            q5 = quantile(diff_time_hrs, probs = 0.05),
+            q80 = quantile(diff_time_hrs, probs = 0.8),
+            q95 = quantile(diff_time_hrs, probs = 0.95)) %>% 
+  ungroup()
+
+  dataset %>% 
+    mutate(col_select = "all_combined") %>% 
     group_by(col_select) %>% 

### 2023-08-12 draft_v0_1.Rmd
Created new draft version for new flow of document. From Windows.

--- 2023-08-12 draft_v0.Rmd
@@
 
+Describe brief, outline of document.
+
 ```{r setup2}
@@
 library(knitr)
+library(lubridate)
 library(tidyverse)
@@
 # colours and theme
+  # COCT colours
 
@@
 
-```{r data_prep}
+## Methods
+
+Describe data prep, code generalisable, tech details - missingness, exclusion decisions, choice of vars (e.g. code var).
+
+```{r data_prep, results='hide'}
+# output from this code chunk should not print in report
 
@@
 
+# data checks ----------------------------------------
+
 glimpse(dta)
 
-# prepare ----------------------------------------------------------
-  # make month and year columns
+sum(is.na(dta$notification_number))
+# no missing notification nrs
+
+length(dta$notification_number) == length(unique(dta$notification_number))
+# each row is unique entry
+
+  # numeric cols ---------
+
+summary(dta$latitude)
+summary(dta$longitude)
+
+  # date cols -----------
+
+unique(lubridate::as_date(dta$creation_timestamp)) 
+# none missing
+# exlude 1 date in dec-2019
+unique(lubridate::as_date(dta$completion_timestamp))
+# some NAs, explore later
+
+# create new variables -----------------------------------------
+
+  # make month and year columns ------
 
@@
 
-unique(dta$month_creation)
 dta$month_creation <- factor(dta$month_creation, levels = 1:12,  labels = mnths)
@@
 
-```
+  # make time-to-complete columns ------
+  # note %--% is a lubridate-specific operator
 
-```{r data_checks}
-sum(is.na(dta$notification_number))
-# no missing notification nrs
-
-length(dta$notification_number) == length(unique(dta$notification_number))
-# each row is unique entry
-
-# missingness
-
-x <- nrow(dta)
-
-map_dbl(dta, ~sum(is.na(.x))/x) %>% 
-  sort()
-
-dta %>% 
-  group_by(is.na(directorate), is.na(department), is.na(branch), 
-           is.na(section), is.na(official_suburb)) %>% 
-  count()
-
-# dates
+dta <- dta %>% 
+   mutate(interv = creation_timestamp %--% completion_timestamp,
+          time_hours = as.duration(interv) / dhours(1),
+          time_days = as.duration(interv) / ddays(1),
+          time_weeks = as.duration(interv) / dweeks(1)) %>% 
+  select(-interv)
 
@@
 
-### Description of the data and preparation steps
-
 #### Missingness
 
-We investigated missingness that would make interpretation hard.
-Entries for which any of the Directorate, Department, Branch, Section or Official suburb were missing are visualised (Figure A).
-Entries with missing Directorate were in general also missing for the other categories.
-Entries for which Directorate was missing, but Official suburb was not(n = 1256), accounted for at most 4% of entries for any suburb and the suburb with the highest percentage of missing entries of this kind, was Kewtown(Figure B).
+```{r missing1}
 
-```{r missing1}
+n_cutoff <- 5
+
+tmp <- dta %>% 
+  group_by(official_suburb) %>% 
+  count(name = "total_suburb") %>% 
+  inner_join(dta %>% 
+  filter(is.na(directorate) & !is.na(official_suburb)) %>% 
+  group_by(official_suburb) %>% 
+  count(name = "n_missing")) %>% 
+  mutate(percentage_missing = n_missing/total_suburb*100) %>% 
+  filter(n_missing >= n_cutoff) %>% 
+  arrange(desc(percentage_missing), desc(n_missing)) %>% 
+  ungroup()
+
+# max(tmp$percentage_missing)
+# tmp$official_suburb[tmp$percentage_missing == max(tmp$percentage_missing)]
+
 tmp <- dta %>% 
@@
   as.data.frame()
+head(tmp)
+
+mis_dir_nonmis_sub <- tmp %>% 
+  group_by(directorate, official_suburb) %>% 
+  count() %>% 
+  filter(directorate == 1 & official_suburb == 0) %>% 
+  pull(n)
+
+```
+
+Missingness of specific data entries can be meaningful, e.g.
+it can point to operational failures if a specific region has more missing data than another.
+Therefore it is important to interrogate patterns of missingness in data.
+We investigated missingness that would make interpretation hard.
+Entries for which any of the Directorate, Department, Branch, Section or Official suburb were missing are visualised (Figure A).
+Entries with missing Directorate were in general also missing for the other categories.
+Suburbs for which there were at least `r n_cutoff` entries, and for which Directorate was missing, but Official suburb was not(n = `r mis_dir_nonmis_sub`), accounted for at most `r tmp$percentage_missing` of entries for any suburb and the suburb with the highest percentage of missing entries of this kind, was `r tmp$official_suburb[tmp$percentage_missing == max(tmp$percentage_missing)]`(Figure B).
+
+```{r missing2}
 
@@
 # missing directorate and dates
-
-unique(dta$month_creation)
 
@@
 ```
+
+```{r}
+
+# prepare ----------------------------------------------------------
+
+  # exclusions -------
+
+# exclude creation date 2019, as only 31 Dec is represented
+
+dta <- dta %>% 
+  filter(creation_timestamp >= lubridate::as_datetime("2020-01-01 00:00:00 UTC"))
+
+```
+
+### Description of the data and preparation steps
 
@@
 
+  sub_txt <- "ALL COMBINED"
+  code_txt <- "All codes combined"
+
   if(is.null(time_cutoff)){
@@
   tabledta <- dataset %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
-  mutate(official_suburb = "ALL COMBINED") %>% 
+  mutate(official_suburb = sub_txt) %>% 
   bind_rows(dataset %>% 
   group_by(official_suburb) %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
     ungroup()) %>% 
-  mutate(code = "all_codes_combined") %>% 
+  mutate(code = code_txt) %>% 
+    bind_rows(dataset %>% 
+                group_by(code) %>% 
+                summarise(n = n(),
+            med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+                 mutate(official_suburb = sub_txt)) %>% 
   bind_rows(dataset %>% 
   group_by(official_suburb, code) %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
@@
   tabledta_below <- dataset_x %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
-  mutate(official_suburb = "ALL COMBINED") %>% 
+  mutate(official_suburb = sub_txt) %>% 
   bind_rows(dataset_x %>% 
   group_by(official_suburb) %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
     ungroup()) %>% 
-  mutate(code = "all_codes_combined") %>% 
+  mutate(code = code_txt) %>% 
+    bind_rows(dataset_x %>% 
+                group_by(code) %>% 
+                summarise(n = n(),
+            med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+              mutate(official_suburb = sub_txt)) %>% 
   bind_rows(dataset_x %>% 
   group_by(official_suburb, code) %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
@@
   tabledta_above <- dataset_x %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
-  mutate(official_suburb = "ALL COMBINED") %>% 
+  mutate(official_suburb = sub_txt) %>% 
   bind_rows(dataset_x %>% 
   group_by(official_suburb) %>% 
-  summarise(med_time_hrs = median(time_hrs),
+  summarise(n = n(),
+            med_time_hrs = median(time_hrs),
             q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
     ungroup()) %>% 
-  mutate(code = "all_codes_combined") %>% 
+  mutate(code = code_txt) %>% 
+    bind_rows(dataset_x %>% 
+                group_by(code) %>% 
+                summarise(n = n(),
+            med_time_hrs = median(time_hrs),
+            q80_time_hrs = quantile(time_hrs, probs = 0.8)) %>% 
+              mutate(official_suburb = sub_txt)) %>% 
   bind_rows(dataset_x %>% 
@@
 
+# look at suburbs with n >= 5
+# all codes combined
+
+n_cutoff <- 5
+sub_txt <- "ALL COMBINED"
+code_txt <- "All codes combined"
+
+time_dta_list$all_times %>% 
+  filter(official_suburb == sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  kable(caption = "City-wide, all codes combined") %>%
+  kable_styling(full_width = FALSE, position = "left")
+
+time_dta_list$all_times %>% 
+  filter(official_suburb != sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  arrange(desc(med_time_hrs)) %>% 
+  slice(1:5) %>% 
+  kable(caption = "Top 5 suburbs with long median completion times, all codes combined") %>%
+  kable_styling(full_width = FALSE, position = "left")
+
+time_dta_list$all_times %>% 
+  filter(official_suburb != sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  arrange(med_time_hrs) %>% 
+  slice(1:5) %>% 
+  kable(caption = "Top 5 suburbs with short median completion times, all codes combined") %>%
+  kable_styling(full_width = FALSE, position = "left")
+
+# split by codes
+# codes with > 5 requests, per suburb
+# show top 5 per code
+
+n_cutoff <- 5
+sub_txt <- "ALL COMBINED"
+code_txt <- "All codes combined"
+
+tmp <- time_dta_list$all_times %>% 
+  filter(official_suburb == sub_txt &
+           code != code_txt &
+           n >= n_cutoff) %>% 
+  arrange(desc(med_time_hrs))
+
+code_txt <- tmp$code[2]
+
+plotdta <- tmp %>% 
+  mutate(set = "City-wide") %>% 
+  filter(code == code_txt) %>% 
+  bind_rows(time_dta_list$all_times %>% 
+  filter(official_suburb != sub_txt &
+           code == code_txt &
+           n != 1) %>% 
+  arrange(desc(med_time_hrs)) %>% 
+  slice(1:5) %>% 
+    mutate(set = "top_long_times")) %>% 
+  bind_rows(time_dta_list$all_times %>% 
+  filter(official_suburb != sub_txt &
+           code == code_txt,
+         n != 1) %>% 
+  arrange(med_time_hrs) %>% 
+  slice(1:5) %>% 
+    mutate(set = "top_short_times")) %>% 
+  pivot_longer(cols = c(med_time_hrs, q80_time_hrs),
+               names_to = "quantile", values_to = "time_hrs")
+
+x <- plotdta %>% 
+  filter(quantile == "med_time_hrs") %>% 
+  select(official_suburb, time_hrs) %>% 
+  arrange(desc(time_hrs)) %>% 
+  pull(official_suburb)
+
+plotdta$official_suburb <- factor(plotdta$official_suburb,
+                                  levels = unique(x))
+
+plotdta %>% 
+  ggplot(aes(time_hrs, official_suburb, colour = set))+
+  geom_point()
+
+# per month
+
 # weekends, public holidays?
@@
 
-# START HERE
 plot_time_per_var <- function(dataset = plotdta, column = "code_group", hours_cutoff = 96){

### 2023-08-12 draft_v0_1.Rmd
Comparisons add, pre-dashboard mock-up. From Windows.

--- 2023-08-11 draft_v0.Rmd
@@
 ---
-title: "Untitled"
-output: html_document
-date: "2023-08-11"
+title: "COCT Urban Mobility"
+output: pdf_document
+author:  "E Maasdorp"
+date: "`r format(Sys.time(), '%d %B %Y')`"
+header-includes:
+  - \usepackage{titling}
+  - \pretitle{\begin{center}
+    \includegraphics[width=2in,height=2in]{img/city_emblem.png}\LARGE\\}
+  - \posttitle{\end{center}}
 editor_options: 
@@
 
+**Note from the author:  This report was compiled to complete the City of Cape Town (COCT) data science challenge.  It is not an official COCT report.**  
+The goal of this analysis was to highlight COCT suburbs that the Urban Mobility Directorate should focus on, for infrastructure improvements, based on service request data for 4 months (December 2020 to March 2021).
+It also explored intervals between creation and completion of service requests, to identify performance across the city.
+Finally, a suggested performance tracking dashboard is provided, and insights are provided to aid
+
 ```{r setup2}
@@
 # colours and theme
+  # COCT colours
+
+coct_cols <- c("#C7006E", "#B9CF00","#00A6CD", "#FF9900",  "#000000")
+
+theme_set(theme_light())
 
@@
 
+# parameters
+
+select_directorate <- "URBAN MOBILITY"
+# specify the Directorate this analysis focuses on
+
 ```
 
-```{r data_prep}
+## Methods
+
+Describe data prep, code generalisable, tech details - missingness, exclusion decisions, choice of vars (e.g. code var).
+
+```{r data_prep, results='hide'}
+# output from this code chunk should not print in report
 
@@
 
-### Description of the data and preparation steps
-
 #### Missingness
 
-We investigated missingness that would make interpretation hard.
-Entries for which any of the Directorate, Department, Branch, Section or Official suburb were missing are visualised (Figure A).
-Entries with missing Directorate were in general also missing for the other categories.
-Entries for which Directorate was missing, but Official suburb was not(n = 1256), accounted for at most 4% of entries for any suburb and the suburb with the highest percentage of missing entries of this kind, was Kewtown(Figure B).
+```{r missing1}
 
-```{r missing1}
+n_cutoff <- 5
+
+tmp1 <- dta %>% 
+  group_by(official_suburb) %>% 
+  count(name = "total_suburb") %>% 
+  inner_join(dta %>% 
+  filter(is.na(directorate) & !is.na(official_suburb)) %>% 
+  group_by(official_suburb) %>% 
+  count(name = "n_missing")) %>% 
+  mutate(percentage_missing = n_missing/total_suburb*100) %>% 
+  filter(n_missing >= n_cutoff) %>% 
+  arrange(desc(percentage_missing), desc(n_missing)) %>% 
+  ungroup()
+
+# max(tmp$percentage_missing)
+# tmp$official_suburb[tmp$percentage_missing == max(tmp$percentage_missing)]
+
 tmp <- dta %>% 
@@
   as.data.frame()
+
+mis_dir_nonmis_sub <- tmp %>% 
+  group_by(directorate, official_suburb) %>% 
+  count() %>% 
+  filter(directorate == 1 & official_suburb == 0) %>% 
+  pull(n)
+
+```
+
+Missingness of specific data entries can be meaningful, e.g.
+it can point to operational failures if a specific region has more missing data than another.
+Therefore it is important to interrogate patterns of missingness in data.
+We investigated missingness that would make interpretation hard.
+Entries for which any of the Directorate, Department, Branch, Section or Official suburb were missing are visualised (Figure A).
+Entries with missing Directorate were in general also missing for the other categories.
+Suburbs for which there were at least `r n_cutoff` entries, and for which Directorate was missing, but Official suburb was not(n = `r mis_dir_nonmis_sub`), accounted for at most `r max(tmp1$percentage_missing)` of entries for any suburb and the suburb with the highest percentage of missing entries of this kind, was `r tmp1$official_suburb[tmp1$percentage_missing == max(tmp1$percentage_missing)]`(Figure B).
+
+```{r missing2}
 
@@
 
+n_cutoff <- 5
 plotdta <- dta %>% 
@@
   mutate(percentage_missing = n_missing/total_suburb*100) %>% 
-  arrange(desc(n_missing), desc(percentage_missing)) %>% 
+  filter(n_missing >= n_cutoff) %>% 
+  arrange(desc(percentage_missing)) %>% 
   ungroup()
@@
   mutate(percentage_missing = n_missing/total_ym*100) %>% 
-  kable() %>%
-  kable_styling(full_width = FALSE, position = "left")
+  kable(format = "simple") %>%
+   kable_styling(position = "left")
 
@@
 
+```
+
+Missingness in specific directorate
+
+```{r}
+
+# Missingness in specific directorate
+
+dta_dir <- dta %>% 
+  filter(directorate == select_directorate)
+
+x <- which(names(dta_dir) %in% c("code_group", "code", "cause_code_group", "cause_code"))
+
+# map(dta_dir[x], ~sort(unique(.x))) 
+# map(dta_dir[x], ~sum(is.na(.x))) 
+# 
+# dta_dir %>%
+#   select(all_of(x)) %>% 
+#   group_by(is.na(.)) %>% 
+#   count()
+
+x <- nrow(dta_dir)
+
+x1 <- dta_dir %>% 
+  group_by(mis_code_grp = is.na(cause_code_group), 
+           mis_cause_code = is.na(cause_code)) %>% 
+  count(name = "missing_cause") %>% 
+  ungroup() %>% 
+  filter(mis_code_grp == TRUE & mis_cause_code == TRUE) %>% 
+  pull(missing_cause)
+
+# tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of {select_directorate} entries.")) %>% 
+#   kable(col.names = NULL) %>%
+#   kable_styling(full_width = FALSE, position = "left")
+
+```
+
+Missing completion dates for `r select_directorate`.
+
+```{r}
+
+dta_dir %>% 
+  group_by(official_suburb) %>% 
+  summarise(n = n(),
+            missing_completion_time = 
+              sum(is.na(completion_timestamp))) %>% 
+  filter(missing_completion_time != 0) %>% 
+  kable(format = "simple") %>%
+  kable_styling(position = "left")
+
+```
+
+```{r}
+
+# prepare ----------------------------------------------------------
+
+  # exclusions -------
+
+# exclude creation date 2019, as only 31 Dec is represented
+
+dta <- dta %>% 
+  filter(creation_timestamp >= lubridate::as_datetime("2020-01-01 00:00:00 UTC"))
+
 # Exclude if missing directorate
@@
 
-# Missingness in urban mobility
-unique(dta$directorate)
+```
 
-dta_um <- dta %>% 
-  filter(directorate == "URBAN MOBILITY")
+## Overview of all Directorates
 
-x <- which(names(dta_um) %in% c("code_group", "code", "cause_code_group", "cause_code"))
-
-map(dta_um[x], ~sort(unique(.x))) 
-map(dta_um[x], ~sum(is.na(.x)))
-
-dta_um %>%
-  select(all_of(x)) %>% 
-  group_by(is.na(.)) %>% 
-  count()
-
-x <- nrow(dta_um)
-
-x1 <- dta_um %>% 
-  group_by(is.na(cause_code_group), is.na(cause_code)) %>% 
-  count(name = "missing_cause") %>% 
-  ungroup() %>% 
-  rename(...1 = 1, ...2 = 2) %>% 
-  filter(...1 == TRUE & ...2 == TRUE) %>% 
-  pull(missing_cause)
-
-tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of Urban Mobility entries.")) %>% 
-  kable(col.names = NULL) %>%
-  kable_styling(full_width = FALSE, position = "left")
-
-```
+Contextualise
 
@@
 
-dta %>% 
+tmp <- dta %>% 
   group_by(directorate) %>% 
@@
   arrange(desc(Nr_of_entries)) %>% 
-  rename(Directorate = directorate) %>% 
-  kable(title = "Number and percentage of entries per Directorate") %>%
-  kable_styling(full_width = FALSE, position = "left")
+  rename(Directorate = directorate)
+
+x <- which(tmp$Directorate == select_directorate)
+
+tmp %>% 
+  kable(format = "simple", caption = "Number and percentage of entries per Directorate") %>%
+  kable_styling( position = "left") %>% 
+  row_spec(row = x, bold = TRUE)
 
@@
   geom_col()+
-  facet_wrap(~year_creation, scales = "free_x") # resize left bar?
+  facet_wrap(~year_creation, scales = "free_x")
+
+tmp <- dta %>% 
+  filter(!is.na(official_suburb)) %>% 
+  group_by(directorate) %>% 
+  count(name = "total_directorate") %>% 
+  inner_join(dta %>% 
+               filter(!is.na(official_suburb)) %>% 
+               group_by(directorate, official_suburb) %>% 
+               count(name = "total_suburb")) %>% 
+  mutate(percentage = total_suburb/total_directorate*100)
+
+map(unique(tmp$directorate),
+    ~tmp %>% 
+      filter(directorate == .x) %>% 
+      arrange(desc(percentage)) %>% 
+      slice(1:3)) %>% 
+  bind_rows() %>% 
+  ggplot(aes(percentage, official_suburb, fill = directorate))+
+  geom_col()+
+  facet_wrap(~directorate, scales = "free_y",
+             nrow = 2)+
+  theme(legend.position = "none")
 
 ```
+
+## Focused Analysis: Directorate `r select_directorate`
 
@@
 ```{r}
-# department, branch, section, code_group not helpful
-dta_um %>% 
-  group_by(code, cause_code) %>% 
-  count() %>% 
-  print(n = Inf)
 
-# filter out cause codes "Not Defined" etc, but keep NA
+# Number of requests over time, including missing suburbs,
+# including missing completion times
 
-dta_um <- dta_um %>% 
-  filter(!cause_code %in% c("No Cause Found",
-                           "None/Private") &
-           !is.na(official_suburb))
+dta_dir %>% 
+  group_by(year_creation, month_creation) %>% 
+  count()
 
-dta_um %>% 
-  group_by(official_suburb) %>% 
-  count() %>% 
-  arrange(desc(n)) %>% 
-  print(n = 20)
+# % suburb missing
 
-# top 3 by month
+dta_dir %>% 
+  group_by(year_creation, month_creation) %>% 
+  count(name = "total_ym") %>% 
+  inner_join(dta_dir %>% 
+               filter(is.na(official_suburb)) %>% 
+               group_by(year_creation, month_creation) %>% 
+               count(name = "total_missing")) %>% 
+  mutate(percentage_missing_suburb = round(total_missing/total_ym*100, 2)) %>% 
+  kable(format = "simple") %>%
+  kable_styling( position = "left")
 
-plotdta <- dta_um %>% 
-  group_by(year_creation, month_creation, official_suburb) %>% 
-  count() %>% 
-  arrange(year_creation, month_creation, desc(n)) %>% 
-  print(n = 20) %>% 
-  ungroup()
+# exclude missing suburb
 
-plotdta <- map(unique(plotdta$month_creation),
-           ~plotdta %>% 
-  filter(month_creation == .x) %>% 
-  arrange(desc(n)) %>% 
-  slice(1:10)) %>% 
+dta_dir <- dta_dir %>% 
+  filter(!is.na(official_suburb))
+
+# nr over time, per suburb, missing suburbs excluded now
+
+tmp <- dta_dir %>% 
+  group_by(year_creation, month_creation) %>% 
+  count(name = "total_ym") %>% 
+  inner_join(dta_dir %>% 
+               group_by(year_creation, 
+                        month_creation, official_suburb) %>% 
+  count(name = "total_suburb")) %>% 
+  mutate(percentage_suburb = total_suburb/total_ym*100) %>% 
+  arrange(year_creation, month_creation,
+          desc(percentage_suburb))
+
+grid <- expand_grid(yr = unique(tmp$year_creation),
+                    mnth = unique(tmp$month_creation))
+
+tmp <- map2(grid$yr, grid$mnth,
+     ~tmp %>% 
+       filter(year_creation == .x & month_creation == .y) %>% 
+       arrange(desc(percentage_suburb)) %>% 
+       slice(1:5)) %>% 
   bind_rows()
 
-dta_um %>% 
+tmp %>% 
+  ggplot(aes(percentage_suburb, reorder(official_suburb, percentage_suburb), fill = month_creation))+
+  geom_col()+
+  facet_wrap(~year_creation + month_creation,
+             scales = "free_y")
+
+# department, branch, section, code_group not helpful
+# dta_dir %>% 
+#   group_by(code, cause_code) %>% 
+#   count() %>% 
+#   print(n = Inf)
+
+# many cause codes NA, use code var only
+
+x <- nrow(dta_dir)
+
+tmp <- dta_dir %>% 
   group_by(official_suburb) %>% 
-  count(name = "total_overall") %>% 
-  inner_join(plotdta) %>% 
-  ggplot(aes(n, official_suburb, fill = total_overall))+
-  geom_col()+
-  facet_wrap(~month_creation)
+  count(name = "total_suburb") %>% 
+  mutate(percentage_suburb = total_suburb/x*100) %>% 
+  arrange(desc(percentage_suburb)) %>% 
+  ungroup() %>% 
+  slice(1:20)
 
-# START HERE
+tmp %>% 
+  kable(format = "simple") %>%
+  kable_styling( position = "left")
+
+#
+
+tmp <- dta_dir %>% 
+  group_by(year_creation, month_creation) %>% 
+  count(name = "total_month") %>% 
+  inner_join(dta_dir %>% 
+  group_by(year_creation, month_creation, official_suburb) %>% 
+  count(name = "total_suburb")) %>% 
+  ungroup() %>% 
+    mutate(percentage = total_suburb/total_month*100) %>% 
+    arrange(year_creation, month_creation, desc(percentage)) %>% 
+  select(year_creation, month_creation, official_suburb,
+         percentage) %>% 
+    pivot_wider(names_from = month_creation, values_from = percentage) %>%  # add median across months
+inner_join(dta_dir %>% 
+             group_by(month_creation) %>% 
+             count(name = "total_month") %>% 
+             inner_join(dta_dir %>% 
+  group_by(month_creation, official_suburb) %>% 
+  count()) %>% 
+  ungroup() %>% 
+    mutate(percentage_suburb = n/total_month*100) %>% 
+    group_by(official_suburb) %>% 
+  summarise(med_percentage = median(percentage_suburb)) %>% 
+  ungroup()) %>% 
+  inner_join(tmp %>% 
+               select(official_suburb)) %>% 
+  arrange(desc(med_percentage))
+
+tmp %>% 
+  kable(format = "simple") %>%
+  kable_styling( position = "left")
+
+# explore codes for top 5
+
+x <- tmp$official_suburb[1:5]
+
+dta_dir %>% 
+  filter(official_suburb %in% x) %>% 
+  group_by(official_suburb, code) %>% 
+  count() %>% 
+  ungroup() %>% 
+  ggplot(aes(n, code, fill = official_suburb))+
+  geom_col(position = "dodge")
+
+# top 3
+
+select_suburbs <- c("PARKLANDS", "MILNERTON")
 
@@
 
-# completion time missing
-x <- nrow(dta_um)
+# exclude data with no completion time, create diff var
 
-dta_um %>% 
-  group_by(is.na(completion_timestamp)) %>% 
-  count(name = "Nr_missing") %>% 
-  ungroup() %>% 
-  rename(Missing_completion_time = 1) %>% 
-  mutate(across(Missing_completion_time, ~ifelse(.x == TRUE, "yes", "no")),
-         Percentage = round(Nr_missing/x*100, 2)) %>% 
-  kable() %>%
-  kable_styling(full_width = FALSE, position = "left")
-
-# breakdown of entries with missing completion times
-
-x <- c("department", "branch", "section", "code_group",
-  "code", "cause_code_group", "cause_code", "official_suburb")
-
-tmp <- map(x, 
-    ~dta_um %>% 
-  filter(is.na(completion_timestamp)) %>% 
-  select(notification_number, all_of(.x)) %>% 
-  mutate(across(-notification_number, ~paste(cur_column(), .x, sep = ":")),
-         value = 1) %>% 
-    rename(key = 2) %>% 
-  pivot_wider(names_from = key, values_from = value)) %>% 
-  purrr::reduce(inner_join) %>% 
-  as.data.frame()
-
-tmp[is.na(tmp)] <- 0
-
-UpSetR::upset(tmp)
-
-# exclude data with no completion time
-
-dta_um <- dta_um %>% 
+dta_dir <- dta_dir %>% 
   filter(!is.na(completion_timestamp))
@@
 
-plotdta <- dta_um %>% 
-  mutate(diff_time_hrs = as.integer(completion_timestamp - creation_timestamp)/60) %>% 
-  select(notification_number, diff_time_hrs, code_group, code, cause_code_group,
-         cause_code, official_suburb)
+# summary(dta_dir$time_hours)
 
-plot_time_per_var <- function(dataset = plotdta, column = "code_group", hours_cutoff = 96){
+# check for outliers in time_hours
 
-  x <- which(names(dataset) == column)
-  dataset <- dataset %>% 
-    rename(col_select = x)
+# tmp <- dta_dir %>%
+#   summarise(q25 = quantile(time_hours, probs = 0.25),
+#             q75 = quantile(time_hours, probs = 0.75),
+#             iqr = q75 - q25,
+#             ol_cutoff = 1.5*iqr + q75)
+# 
+# dta_dir %>% 
+#   group_by(time_hours > tmp$ol_cutoff) %>% 
+#   count()
+# 
+# dta_dir %>% 
+#   filter(time_days >= tmp$ol_cutoff) %>% 
+#   ggplot(aes(time_hours, official_suburb, 
+#              colour = official_suburb))+
+#   geom_jitter()+
+#   facet_wrap(~code, scales = "free_y")+
+#   theme(legend.position = "none")
 
-  tmp <- dataset %>% 
-  group_by(col_select) %>% 
+#  create table with medians, q80 for all suburbs and for all of city
+
+time_tables_make <- function(dataset = dta_dir, time_cutoff = NULL){
+
+  sub_txt <- "ALL COMBINED"
+  code_txt <- "All codes combined"
+
+  if(is.null(time_cutoff)){
+
+  tabledta <- dataset %>% 
   summarise(n = n(),
-            med_hrs = median(diff_time_hrs),
-            q5 = quantile(diff_time_hrs, probs = 0.05),
-            q80 = quantile(diff_time_hrs, probs = 0.8),
-            q95 = quantile(diff_time_hrs, probs = 0.95)) %>% 
-  ungroup()
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+  mutate(official_suburb = sub_txt) %>% 
+  bind_rows(dataset %>% 
+  group_by(official_suburb) %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+    ungroup()) %>% 
+  mutate(code = code_txt) %>% 
+    bind_rows(dataset %>% 
+                group_by(code) %>% 
+                summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+                 mutate(official_suburb = sub_txt)) %>% 
+  bind_rows(dataset %>% 
+  group_by(official_suburb, code) %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+    ungroup())
 
-p1 <- tmp %>% 
-  filter(q95 < hours_cutoff) %>% 
-  select(col_select) %>% 
-  inner_join(dataset %>% 
-               select(col_select, diff_time_hrs)) %>% 
-  ggplot(aes(diff_time_hrs, col_select))+
-  ggdist::stat_interval()+
-  labs(y = column)
+  return(tabledta)
 
-p2 <- tmp %>% 
-  filter(q5 >= hours_cutoff) %>% 
-  select(col_select) %>% 
-  inner_join(dataset) %>% 
-  mutate(time_weeks = (diff_time_hrs/24)/7) %>% 
-  ggplot(aes(time_weeks, col_select))+
-  ggdist::stat_interval()+
-  labs(y = column)
+} else if(!is.null(time_cutoff)){
 
-return(list(hours = p1, weeks = p2))
+  dataset_x <- dataset %>% 
+    filter(time_hours < time_cutoff)
 
+  tabledta_below <- dataset_x %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+  mutate(official_suburb = sub_txt) %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb) %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+    ungroup()) %>% 
+  mutate(code = code_txt) %>% 
+    bind_rows(dataset_x %>% 
+                group_by(code) %>% 
+                summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+              mutate(official_suburb = sub_txt)) %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb, code) %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+    ungroup())
+
+  dataset_x <- dataset %>% 
+    filter(time_hours >= time_cutoff)
+
+  tabledta_above <- dataset_x %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+  mutate(official_suburb = sub_txt) %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb) %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+    ungroup()) %>% 
+  mutate(code = code_txt) %>% 
+    bind_rows(dataset_x %>% 
+                group_by(code) %>% 
+                summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+              mutate(official_suburb = sub_txt)) %>% 
+  bind_rows(dataset_x %>% 
+  group_by(official_suburb, code) %>% 
+  summarise(n = n(),
+            med_time_hours = median(time_hours),
+            q80_time_hours = quantile(time_hours, probs = 0.8)) %>% 
+    ungroup())
+
+  return(list(tabledta_below = tabledta_below,
+                tabledta_above = tabledta_above))
+}
 }
 
-plotlist <- plot_time_per_var(dataset = plotdta, column = "code_group", hours_cutoff = 96)
+time_dta_list <- list()
 
-ggpubr::ggarrange(plotlist = plotlist)
+# for all of city, all codes, all completion time lengths
 
-plotlist <- plot_time_per_var(dataset = plotdta, column = "code", hours_cutoff = 96)
+time_dta_list$all_times <- time_tables_make(dataset = dta_dir, time_cutoff = NULL)
 
-ggpubr::ggarrange(plotlist = plotlist)
+# for all of city, all codes, split by times < or > 24hrs
+tmp <- time_tables_make(dataset = dta_dir, time_cutoff = 24)
 
-plotlist <- plot_time_per_var(dataset = plotdta, column = "cause_code_group", hours_cutoff = 96)
+time_dta_list$below_24 <- tmp$tabledta_below 
+time_dta_list$above_eq_24 <- tmp$tabledta_above
 
-ggpubr::ggarrange(plotlist = plotlist)
+# look at suburbs with n >= 5
+# all codes combined
 
-plotlist <- plot_time_per_var(dataset = plotdta, column = "cause_code", hours_cutoff = 96)
+n_cutoff <- 5
+sub_txt <- "ALL COMBINED"
+code_txt <- "All codes combined"
 
-ggpubr::ggarrange(plotlist = plotlist)
+time_dta_list$all_times %>% 
+  filter(official_suburb == sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  kable(format = "simple", caption = "City-wide, all codes combined") %>%
+  kable_styling( position = "left")
 
-plotdta %>% 
-  ggplot(aes(diff_time_hrs, official_suburb))+
-  ggdist::stat_interval()
+# filter by < 24 hrs
 
-ggpubr::ggarrange(plotlist = plotlist)
+time_dta_list$below_24 %>% 
+  filter(official_suburb == sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in less than 24 hours") %>%
+  kable_styling( position = "left")
+
+# > 24 hrs
+
+time_dta_list$above_eq_24 %>% 
+  filter(official_suburb == sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in 24 hours or more") %>%
+  kable_styling( position = "left")
+
+# split by 24 hours
+
+x <- time_dta_list$below_24 %>% 
+  filter(official_suburb == "ALL COMBINED" &
+           code == "All codes combined") %>% 
+  pull(n)
+
+x <- time_dta_list$all_times %>% 
+  filter(official_suburb == "ALL COMBINED" &
+           code == "All codes combined") %>% 
+  mutate(perc = x/n*100) %>% 
+  pull(perc)
+
+# x
+
+time_dta_list$all_times %>% 
+  filter(official_suburb != sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  arrange(desc(med_time_hours)) %>% 
+  slice(1:5) %>% 
+  kable(format = "simple", caption = "Top 5 suburbs with longest median completion times, all codes combined") %>%
+  kable_styling( position = "left")
+
+time_dta_list$all_times %>% 
+  filter(official_suburb != sub_txt &
+           code == code_txt &
+           n >= n_cutoff) %>% 
+  arrange(med_time_hours) %>% 
+  slice(1:5) %>% 
+  kable(format = "simple", caption = "Top 5 suburbs with shortest median completion times, all codes combined") %>%
+  kable_styling(position = "left")
+
+# split by codes
+# codes with > 5 requests, per suburb
+# show top 5 per code
+
+# get codes for which > 5 requests for some suburbs
+
+n_cutoff <- 5
+
+x <- dta_dir %>% 
+  group_by(code, official_suburb) %>% 
+  count() %>% 
+  arrange(desc(n)) %>% 
+  filter(n >= n_cutoff) %>% 
+  ungroup() %>% 
+  distinct(code) %>% 
+  pull(code)
+
+sub_txt <- "ALL COMBINED"
+time_cutoff <- 24
+
+p1 <- map(x,
+    ~time_dta_list$below_24 %>% 
+  filter(official_suburb != sub_txt &
+           code == .x &
+           n >= n_cutoff) %>% 
+  select(official_suburb, code) %>% 
+    distinct() %>% 
+    inner_join(dta_dir %>% 
+                 filter(time_hours < time_cutoff))) %>% 
+  bind_rows() %>% 
+  ggplot(aes(time_hours, official_suburb))+
+    ggdist::stat_interval()+
+  facet_wrap(~code, scales = "free_y", ncol = 1)
+
+p2 <- map(x,
+    ~time_dta_list$above_eq_24 %>% 
+  filter(official_suburb != sub_txt &
+           code == .x &
+           n >= n_cutoff) %>% 
+  select(official_suburb, code) %>% 
+    distinct() %>% 
+    inner_join(dta_dir %>% 
+                 filter(time_hours >= time_cutoff))) %>% 
+  bind_rows() %>% 
+  ggplot(aes(time_hours, official_suburb))+
+    ggdist::stat_interval()+
+  facet_wrap(~code, scales = "free_y",
+             ncol = 1)
+
+ggpubr::ggarrange(p1, p2, ncol = 1, nrow = 2,
+                  heights = c(2,1))
+
+tmp <- map(x,
+           ~time_dta_list$all_times %>% 
+  filter(official_suburb != sub_txt &
+           code == .x &
+           n >= n_cutoff) %>% 
+  arrange(desc(med_time_hours)))
+
+x <- seq(1,length(tmp), by = 2)
+
+kable(tmp[1:2], format = "simple") %>%
+    kable_styling( position = "left")
+
+# per month
+
+# weekends, public holidays?
 
@@
 
+```{r}
+
+time_dta_list$all_times %>% 
+  filter(official_suburb %in% c(select_suburbs, "ALL COMBINED"))
+
+# wilcox test
+map(select_suburbs,
+    ~dta_dir %>% 
+  select(official_suburb, time_hours) %>%
+  mutate(suburb_bin = case_when(official_suburb %in% 
+                                  .x ~ .x,
+                                TRUE ~ "ALL COMBINED")) %>% 
+  summarise(suburb = .x,
+            p_value = wilcox.test(time_hours ~ 
+                                    suburb_bin)$p.value))
+
+# prop.test
+
+prop_test_list <- list()
+
+for(i in 1:length(select_suburbs)){
+x <- time_dta_list$all_times %>% 
+  filter(official_suburb %in% "ALL COMBINED" &
+           code == "All codes combined") %>% 
+  pull(med_time_hours)
+
+tmp <- dta_dir %>% 
+  select(official_suburb, time_hours) %>%
+  mutate(med_compare = case_when(time_hours < x ~ "below_coct_median",
+                                time_hours >= x ~ "above_or_eq_coct_med")) %>% 
+  mutate(suburb_bin = case_when(official_suburb %in% 
+                                  select_suburbs[i] ~ select_suburbs[i],
+                                TRUE ~ "ALL COMBINED")) %>% 
+  group_by(med_compare, suburb_bin) %>% 
+  count() %>% 
+  pivot_wider(names_from = med_compare, values_from = n) %>% 
+  ungroup() %>% 
+  select(-suburb_bin)
+
+prop_test_list[[i]] <- tibble(suburb = select_suburbs[i],
+       p_value = prop.test(as.matrix(tmp))$p.value)
+
+}
+
+prop_test_list
+
+```
+
 ## 3.
@@
 ## 4. Identify value-adding insights for the management of Urban Mobility, from the dataset provided, in regard to commuter transport within the City.
+
+```{r}
+# traffic congestion over time
+
+ unique(dta_dir$code)
+
+ code_txt <- "Conjested Traffic at Intersection"
+
+tmp <- dta_dir %>% 
+ mutate(across(creation_timestamp, ~as_date(.x))) %>% 
+ filter(code == code_txt) %>% 
+ group_by(code,creation_timestamp) %>% 
+ count() %>% 
+ ungroup()
+
+x <- sort(tmp$n, decreasing = TRUE)
+x <- x[6]
+
+tmp %>% 
+mutate(day_of_week = wday(creation_timestamp, label=TRUE),
+       label = case_when(n > x ~ 
+                           paste(day_of_week, creation_timestamp),
+                         TRUE ~ "")) %>%
+
+ ggplot(aes(creation_timestamp, n, group = code,
+            label = label))+
+ geom_line(colour = "grey70")  +
+ geom_point()+
+   geom_text()
+
+dta_dir %>% 
+ mutate(create_date = as_date(creation_timestamp)) %>% 
+ filter(code == code_txt) %>% 
+ group_by(code,create_date) %>% 
+ count() %>% 
+ ungroup() %>% 
+  filter(n > x) %>% 
+  select(-n) %>% 
+  inner_join(dta_dir) %>% 
+  group_by(create_date,official_suburb) %>% 
+  count() %>% 
+  ungroup() %>% 
+  ggplot(aes(n, reorder(official_suburb,n), fill = as.factor(create_date)))+
+  geom_col()
+
+ # which days of the week have more traffic congestion reports
+
+ x <- dta_dir %>% 
+   group_by(code) %>% 
+   count() %>% 
+   arrange(desc(n)) %>% 
+   ungroup() %>% 
+   slice(1:5) %>% 
+   pull(code)
+
+ dta_dir %>% 
+   filter(code %in% x) %>% 
+ mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
+ group_by(code, day_of_week) %>% 
+ count() %>%
+   ggplot(aes(day_of_week, n, group = code, colour = code))+
+ geom_point()+
+ geom_line()
+
+ dta_dir %>% 
+ mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
+ filter(code == code_txt) %>% 
+ group_by(day_of_week) %>% 
+ count() %>% 
+ ggplot(aes(day_of_week, n))+
+ geom_col()
+
+ # ?stats
+ # odds of having congestion on a wednesday vs other weekdays
+
+ # suburbs and days of week
+
+x <-  dta_dir %>% 
+ mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
+ group_by(day_of_week, official_suburb) %>% 
+ count() %>% 
+   ungroup() %>% 
+   arrange(desc(n)) %>% 
+   slice(1:10) %>% 
+   pull(n)
+
+x <- x[length(x)]
+
+sz <- 3
+ dta_dir %>% 
+ mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
+ group_by(day_of_week, official_suburb) %>% 
+ count() %>% 
+   mutate(label = case_when(n >= x ~ official_suburb,
+                          TRUE ~ "")) %>% 
+   ggplot(aes(day_of_week, n, label = label))+
+   geom_point(alpha = 0.6)+
+   geom_text(alpha = 0.5, size = sz, check_overlap = TRUE)
+
+```

### 2023-08-13 draft_v0_1.Rmd
Text added to report.  Appendix added for tech details. From Windows.

@@
 **Note from the author:  This report was compiled to complete the City of Cape Town (COCT) data science challenge.  It is not an official COCT report.**  
-The goal of this analysis was to highlight COCT suburbs that the Urban Mobility Directorate should focus on, for infrastructure improvements, based on service request data for 4 months (December 2020 to March 2021).
-It also explored intervals between creation and completion of service requests, to identify performance across the city.
-Finally, a suggested performance tracking dashboard is provided, and insights are provided to aid
+The goal of this analysis is to highlight COCT suburbs that the Urban Mobility Directorate should focus on, for infrastructure improvements, based on service request data for 4 months (December 2019 to March 2020).
+It also explored intervals between creation and completion of service requests, to assess performance across the city.
+In future, the identification of performance pain points can be aided by a performance tracking dashboard, for which a visual mock-up is provided.
+Finally, some insights relevant to commuter transport, gained from the current data, are provided.
 
@@
 
-Describe data prep, code generalisable, tech details - missingness, exclusion decisions, choice of vars (e.g. code var).
-
-```{r data_prep, results='hide'}
+The ```sr_hex_truncated.csv``` dataset was provided for this analysis.
+It originates from service requests across COCT spanning December 2019 to March 2020.
+The analysis was performed in the Statistical Programming Language R, using a RMarkdown file to produce this document.
+The code is available in a Github repository () and this analysis should be fully reproducible.
+The R and package versions used are listed at the end of this document in an appendix.
+The dataset was explored and new variables created from existing ones, as necessary.
+Tables and plots were used to communicate useful summaries.
+Medians, quantiles and non-parametric tests for differences in service request completion times were used because of skew data distribution.
+Only 1 day of 2019 is represented in the data (31 December) and it was excluded from the analysis.
+Missing data entries were explored and details focusing on missingness can be found in the appendix at the end of this document.
+Service request data that are hard to interpret because of missing entries in specific variables, were excluded:  rows with missing Directorate for all analyses, rows with missing Official Suburb, for analyses specifically comparing suburbs, and rows with missing service request completion dates for completion time analyses.
+The data contains 4 service request code variables that provide descriptions of the type of request.
+Due to severe missingness in two of these, and the differences in granularity in the remaining two, only the ```code``` variable was used to compare and provide insight into the types and frequency of requests.
+
+```{r data_check, results='hide'}
 # output from this code chunk should not print in report
@@
 
-#### Missingness
-
 ```{r missing1}
 
+#### Missingness - include these results in appendix
+# save in appendix list
+appx_list <- list()
+
 n_cutoff <- 5
 
-tmp1 <- dta %>% 
+appx_list$miss_dir <- dta %>% 
   group_by(official_suburb) %>% 
@@
 
-tmp <- dta %>% 
+appx_list$upset_data <- dta %>% 
   select(directorate, department, branch, section, official_suburb) %>% 
@@
 
-mis_dir_nonmis_sub <- tmp %>% 
+appx_list$mis_dir_nonmis_sub <- appx_list$upset_data %>% 
   group_by(directorate, official_suburb) %>% 
@@
 
-Missingness of specific data entries can be meaningful, e.g.
-it can point to operational failures if a specific region has more missing data than another.
-Therefore it is important to interrogate patterns of missingness in data.
-We investigated missingness that would make interpretation hard.
-Entries for which any of the Directorate, Department, Branch, Section or Official suburb were missing are visualised (Figure A).
-Entries with missing Directorate were in general also missing for the other categories.
-Suburbs for which there were at least `r n_cutoff` entries, and for which Directorate was missing, but Official suburb was not(n = `r mis_dir_nonmis_sub`), accounted for at most `r max(tmp1$percentage_missing)` of entries for any suburb and the suburb with the highest percentage of missing entries of this kind, was `r tmp1$official_suburb[tmp1$percentage_missing == max(tmp1$percentage_missing)]`(Figure B).
-
 ```{r missing2}
 
-na_interpret_affected <- UpSetR::upset(tmp)
-
-na_interpret_affected
-grid::grid.text("Fig. A",x = 0.65, y=0.95, gp=grid::gpar(fontsize=20))
+appx_list$na_interpret_affected <- UpSetR::upset(appx_list$upset_data)
 
@@
 
-plotdta %>% 
+appx_list$plot_mis_dir_nonmis_sub <- plotdta %>% 
   slice(1:20) %>% 
@@
 
-dta %>% 
+appx_list$table_mis_dir_dates <- dta %>% 
   group_by(year_creation, month_creation) %>% 
@@
 
-Of the entries with Official suburb missing, ...
-
-```{r}
+```{r missing3}
+# Of the entries with Official suburb missing, ...
 
@@
 
-plotdta %>% 
+appx_list$mis_sub <- plotdta %>% 
   ggplot()+
@@
 
-Missingness in specific directorate
-
-```{r}
+```{r missing4}
 
@@
 
-# tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of {select_directorate} entries.")) %>% 
-#   kable(col.names = NULL) %>%
-#   kable_styling(full_width = FALSE, position = "left")
-
-```
-
-Missing completion dates for `r select_directorate`.
-
-```{r}
-
-dta_dir %>% 
+appx_list$miss_cause <- tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of {select_directorate} entries.")) %>%
+  kable(col.names = NULL) %>%
+  kable_styling(position = "left")
+
+```
+
+```{r missing5}
+# Missing completion dates for `r select_directorate`.
+appx_list$miss_compl_dates <- dta_dir %>% 
   group_by(official_suburb) %>% 
@@
 
-```{r}
+```{r prep}
 
@@
 
-Contextualise
+This section provides broad context of service requests across all COCT Directorates for the time period included in the dataset.
+The table displays number of service requests per Directorate, as well as the percentage each represents of the total requests for all Directorates.
+The Urban Mobility Directorate requests make up a small percentage of total requests.
 
@@
 
+```
+
+The percentages of total requests seem to remain fairly stable per month, in the included time period, with Water and Sanitation and Energy contributing the bulk of requests.
+
+```{r}
 dta %>% 
@@
   geom_col()+
-  facet_wrap(~year_creation, scales = "free_x")
+  facet_wrap(~year_creation, scales = "free_x")+
+  labs(fill = "Directorate, ordered from smallest to largest contribution",
+       x = "Month", y = "Percentage of total requests")
 
@@
 
+```
+
+Highlighting the top 3 suburbs (largest percentage of total per suburb) for each Directorate, reveals that for most Directorates the total requests are made up of many suburbs, contributing relatively low proportions to the total.
+For Urban Mobility, for example, the biggest contribution is Milnerton, around 4%, which means that many other suburbs contribute percentages lower than 4%, to make up the total of requests for this Directorate.
+The only exception is The Human Settlements Directorate where the top 3 suburbs, together, contribute around 30% of the total.
+
+```{r}
 map(unique(tmp$directorate),
@@
 
-## "In which 3 suburbs should the Urban Mobility directorate concentrate their infrastructure improvement efforts?"
-
-```{r}
+This section provides insight to inform decisions for infrastructure improvement efforts.
+It is important to note that the suburb information was missing for a large percentage of service requests (Table x).
+
+```{r focused1}
 
@@
 
-dta_dir %>% 
-  group_by(year_creation, month_creation) %>% 
-  count()
+# dta_dir %>% 
+#   group_by(year_creation, month_creation) %>% 
+#   count()
 
@@
 
+```
+
+The reason for the suburb missingness is not clear from the data.
+For the top 5 codes with most service requests, "Pothole&Defect Road Foot Bic Way/Kerbs" and "Paint Markings Lines&Signs" were more likely to be missing, than not (Table x).
+
+```{r focused2}
+dta_dir %>% 
+  group_by(code) %>% 
+  count() %>% 
+  ungroup() %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5) %>% 
+  select(-n) %>% 
+  inner_join(dta_dir %>% 
+  group_by(mis_sub = is.na(official_suburb), code) %>% 
+  count()) %>% 
+  pivot_wider(names_from = mis_sub, values_from = n) %>% 
+  print(n = Inf)
+
+```
+
+The odds of a suburb entry being missing was more than double for a number of dates, the top 4 of which occur closely together in March 2020 (Table x).
+There may therefore be a systematic reason for this missingness, but it is not clear from this dataset.
+
+```{r focused5}
+
+dta_dir %>% 
+  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
+  group_by(creation_timestamp, mis_sub = is.na(official_suburb)) %>% 
+  count() %>% 
+  ungroup() %>% 
+  pivot_wider(names_from = mis_sub, values_from = n) %>% 
+  rename(missing = `TRUE`, not_missing = `FALSE`) %>% 
+  mutate(odds = round(missing/not_missing, 2)) %>% 
+  arrange(desc(odds)) %>% 
+  slice(1:5) %>% 
+  kable(format = "simple") %>%
+  kable_styling( position = "left")
+
+tmp <- dta_dir %>% 
+  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
+  group_by(creation_timestamp, mis_sub = is.na(official_suburb)) %>% 
+  count() %>% 
+  ungroup() %>% 
+  filter(mis_sub == TRUE) %>% 
+  arrange(desc(n))
+
+x <- sum(tmp$n)
+
+x <- tmp %>% 
+  slice(1:5) %>% 
+  summarise(perc = sum(n)/x*100)
+
+```
+
+The top 5 dates for missing suburb entries, represent `r x`% of all the missing suburb entries.
+
+Service requests with missing suburb entries were excluded for the rest of this analysis, but it should be kept in mind that a systematic cause for the missingness may influence the interpretation of these results.
+An important action point may therefore be to attempt to find the cause for the large proportion of missing suburb entries.
+
+```{r focused4}
+
+# START HERE
 # exclude missing suburb
@@
 ```
+
+## Appendix
+
+```{r}
+
+#### Missingness - include these results in appendix
+# save in appendix list
+appx_list <- list()
+
+n_cutoff <- 5
+
+appx_list$miss_dir <- dta %>% 
+  group_by(official_suburb) %>% 
+  count(name = "total_suburb") %>% 
+  inner_join(dta %>% 
+  filter(is.na(directorate) & !is.na(official_suburb)) %>% 
+  group_by(official_suburb) %>% 
+  count(name = "n_missing")) %>% 
+  mutate(percentage_missing = n_missing/total_suburb*100) %>% 
+  filter(n_missing >= n_cutoff) %>% 
+  arrange(desc(percentage_missing), desc(n_missing)) %>% 
+  ungroup()
+
+# max(tmp$percentage_missing)
+# tmp$official_suburb[tmp$percentage_missing == max(tmp$percentage_missing)]
+
+appx_list$upset_data <- dta %>% 
+  select(directorate, department, branch, section, official_suburb) %>% 
+  mutate(across(everything(), ~as.numeric(is.na(.x)))) %>% 
+  as.data.frame()
+
+appx_list$mis_dir_nonmis_sub <- appx_list$upset_data %>% 
+  group_by(directorate, official_suburb) %>% 
+  count() %>% 
+  filter(directorate == 1 & official_suburb == 0) %>% 
+  pull(n)
+
+```
+
+Missingness of specific data entries can be meaningful, e.g.
+it can point to operational failures if a specific region has more missing data than another.
+Therefore it is important to interrogate patterns of missingness in data.
+We investigated missingness that would make interpretation hard.
+Entries for which any of the Directorate, Department, Branch, Section or Official suburb were missing are visualised (Figure A).
+Entries with missing Directorate were in general also missing for the other categories.
+Suburbs for which there were at least `r n_cutoff` entries, and for which Directorate was missing, but Official suburb was not(n = `r mis_dir_nonmis_sub`), accounted for at most `r max(tmp1$percentage_missing)` of entries for any suburb and the suburb with the highest percentage of missing entries of this kind, was `r tmp1$official_suburb[tmp1$percentage_missing == max(tmp1$percentage_missing)]`(Figure B).
+
+```{r}
+
+appx_list$na_interpret_affected <- UpSetR::upset(tmp)
+
+na_interpret_affected
+grid::grid.text("Fig. A",x = 0.65, y=0.95, gp=grid::gpar(fontsize=20))
+
+n_cutoff <- 5
+plotdta <- dta %>% 
+  group_by(official_suburb) %>% 
+  count(name = "total_suburb") %>% 
+  inner_join(dta %>% 
+  filter(is.na(directorate) & !is.na(official_suburb)) %>% 
+  group_by(official_suburb) %>% 
+  count(name = "n_missing")) %>% 
+  mutate(percentage_missing = n_missing/total_suburb*100) %>% 
+  filter(n_missing >= n_cutoff) %>% 
+  arrange(desc(percentage_missing)) %>% 
+  ungroup()
+
+appx_list$plot_mis_dir_nonmis_sub <- plotdta %>% 
+  slice(1:20) %>% 
+  ggplot()+
+  geom_point(aes(percentage_missing, reorder(official_suburb, percentage_missing), size = n_missing))+
+  labs(title = "Fig. B: Top 20 Suburbs with missing Directorate")
+
+# missing directorate and dates
+
+appx_list$table_mis_dir_dates <- dta %>% 
+  group_by(year_creation, month_creation) %>% 
+  count(name = "total_ym") %>% 
+  inner_join(dta %>% 
+               group_by(year_creation, month_creation, is.na(directorate)) %>% 
+               count(name = "n_missing")) %>%
+  ungroup() %>% 
+  rename(missing_directorate = 4) %>% 
+  filter(missing_directorate == TRUE) %>% 
+  mutate(percentage_missing = n_missing/total_ym*100) %>% 
+  kable(format = "simple") %>%
+   kable_styling(position = "left")
+
+```
+
+Of the entries with Official suburb missing, ...
+
+```{r}
+
+plotdta <- dta %>% 
+  group_by(directorate) %>% 
+  count(name = "total_directorate") %>%
+  inner_join(dta %>% 
+               filter(is.na(official_suburb)) %>% 
+               group_by(directorate) %>% 
+                        count(name = "n_missing")) %>% 
+  ungroup() %>% 
+  mutate(percentage_missing = n_missing/total_directorate*100)
+
+appx_list$mis_sub <- plotdta %>% 
+  ggplot()+
+  geom_point(aes(percentage_missing, reorder(directorate, percentage_missing), size = n_missing))+
+  labs(title = "Fig. C: Missing Suburbs per Directorate")
+
+```
+
+Missingness in specific directorate
+
+```{r}
+
+# Missingness in specific directorate
+
+dta_dir <- dta %>% 
+  filter(directorate == select_directorate)
+
+x <- which(names(dta_dir) %in% c("code_group", "code", "cause_code_group", "cause_code"))
+
+# map(dta_dir[x], ~sort(unique(.x))) 
+# map(dta_dir[x], ~sum(is.na(.x))) 
+# 
+# dta_dir %>%
+#   select(all_of(x)) %>% 
+#   group_by(is.na(.)) %>% 
+#   count()
+
+x <- nrow(dta_dir)
+
+x1 <- dta_dir %>% 
+  group_by(mis_code_grp = is.na(cause_code_group), 
+           mis_cause_code = is.na(cause_code)) %>% 
+  count(name = "missing_cause") %>% 
+  ungroup() %>% 
+  filter(mis_code_grp == TRUE & mis_cause_code == TRUE) %>% 
+  pull(missing_cause)
+
+appx_list$miss_cause <- tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of {select_directorate} entries.")) %>%
+  kable(col.names = NULL) %>%
+  kable_styling(position = "left")
+
+```
+
+Missing completion dates for `r select_directorate`.
+
+```{r}
+
+appx_list$miss_compl_dates <- dta_dir %>% 
+  group_by(official_suburb) %>% 
+  summarise(n = n(),
+            missing_completion_time = 
+              sum(is.na(completion_timestamp))) %>% 
+  filter(missing_completion_time != 0) %>% 
+  kable(format = "simple") %>%
+  kable_styling(position = "left")
+
+```
+
+```{r}
+
+sessionInfo()
+
+```

### 2023-08-13 draft_v0_1.Rmd
More text added. Plots tidied. From Windows.

@@
 
-theme_set(theme_light())
-
-# functions
+# ggplot theme
+
+theme_doc <- theme_light() + 
+  theme(text = element_text(color = "grey30",
+                            size = 18),
+        line = element_line(color = "grey30"),
+        rect = element_rect(color = "grey30",
+                            fill = "#FFFFF6"),
+        panel.background = element_rect(fill = "#FFFFF6", colour = "grey70"),
+        # axis.line = element_blank(),
+        # axis.ticks = element_blank(),
+        strip.background = element_rect(colour = "grey70",
+                                        fill = "#FFFFF6"),
+        strip.text = element_text(colour = "grey30"))
+
+theme_set(theme_doc)
+# theme_set(theme_light())
 
@@
   ggplot()+
-  geom_point(aes(percentage_missing, reorder(official_suburb, percentage_missing), size = n_missing))+
+  geom_point(aes(percentage_missing, reorder(official_suburb, percentage_missing), size = n_missing), colour = coct_cols[3])+
   labs(title = "Fig. B: Top 20 Suburbs with missing Directorate")
@@
   ggplot()+
-  geom_point(aes(percentage_missing, reorder(directorate, percentage_missing), size = n_missing))+
+  geom_point(aes(percentage_missing, reorder(directorate, percentage_missing), size = n_missing),
+             colour = coct_cols[2])+
   labs(title = "Fig. C: Missing Suburbs per Directorate")
@@
 
-```{r}
+```{r fig.height=12, fig.width=24}
 map(unique(tmp$directorate),
@@
              nrow = 2)+
-  theme(legend.position = "none")
+  scale_fill_manual(values = c(coct_cols, coct_cols))+
+  labs(x = "Percentage")+
+  theme(legend.position = "none",
+        axis.title.y = element_blank())
 
@@
   pivot_wider(names_from = mis_sub, values_from = n) %>% 
-  print(n = Inf)
+  rename(not_missing = `FALSE`, missing = `TRUE`) %>% 
+  kable(format = "simple")
 
@@
 The top 5 dates for missing suburb entries, represent `r x`% of all the missing suburb entries.  
-
 Service requests with missing suburb entries were excluded for the rest of this analysis, but it should be kept in mind that a systematic cause for the missingness may influence the interpretation of these results.
 An important action point may therefore be to attempt to find the cause for the large proportion of missing suburb entries.
+To identify the top suburbs to focus infrastructure improvements on, one can take different approaches.  Simply taking the total number of service requests, would highlight the suburbs in Table x.
 
@@
 
-# START HERE
 # exclude missing suburb
@@
 
+ # simple totals
+tmp <- dta_dir %>% 
+  group_by(official_suburb) %>% 
+  count() %>% 
+  ungroup() %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5)
+
+tmp %>% 
+  kable(format = "simple") %>% 
+  kable_styling(position = "left")
+
+top5 <- tmp %>% pull(official_suburb)
+```
+
+One can also look at consistent numbers of requests over the 3 month period(Fig.
+x).
+Of the top 5 by simple totals, Milnerton, Marconi Beam and Table View are in the top 5 per month, three, two and two times, respectively, while Parklands and Brackenfell Common appear only once.
+
+```{r}
 # nr over time, per suburb, missing suburbs excluded now
+
+# 
+# dta_dir %>% 
+#   filter(official_suburb %in% top5) %>% 
+#   mutate(request_date = as_date(creation_timestamp)) %>% 
+#   group_by(request_date, official_suburb) %>% 
+#   count() %>% 
+#   ungroup() %>% 
+#   filter(official_suburb %in% top5) %>% 
+#   ggplot(aes(official_suburb,
+#              n, colour = month(request_date))) + 
+#   geom_boxplot()
 
@@
   facet_wrap(~year_creation + month_creation,
-             scales = "free_y")
-
-# department, branch, section, code_group not helpful
-# dta_dir %>% 
-#   group_by(code, cause_code) %>% 
-#   count() %>% 
-#   print(n = Inf)
-
-# many cause codes NA, use code var only
-
-x <- nrow(dta_dir)
-
-tmp <- dta_dir %>% 
-  group_by(official_suburb) %>% 
-  count(name = "total_suburb") %>% 
-  mutate(percentage_suburb = total_suburb/x*100) %>% 
-  arrange(desc(percentage_suburb)) %>% 
-  ungroup() %>% 
-  slice(1:20)
-
-tmp %>% 
-  kable(format = "simple") %>%
-  kable_styling( position = "left")
-
-#
-
+             scales = "free_y")+
+  labs(x = "Percentage")+
+  theme(legend.position = "none",
+        axis.title.y = element_blank())+
+  scale_fill_manual(values = coct_cols)
+
+```
+
+Calculating the percentage of requests each suburb represents, per month, and taking the median of those percentages over the three included months, produces Table x.
+
+```{r}
 tmp <- dta_dir %>% 
@@
                select(official_suburb)) %>% 
+  distinct() %>% 
   arrange(desc(med_percentage))
@@
 
-# explore codes for top 5
-
-x <- tmp$official_suburb[1:5]
+```
+
+It is plausible that Camps Bay / Bakoven shows up mainly because of holiday traffic, and isn't routinely a problem area.
+It can also be seen that Milnerton was responsible for an unusually high percentage in February, and Marconi Beam in March.
+Exploring request numbers at a more granular level, reveals that 4 particular dates have a disproportionate influence on the high request numbers in Brackenfell Common, Milnerton and Parklands.
+
+```{r}
+
+tmp <- dta_dir %>% 
+  filter(official_suburb %in% top5) %>% 
+  mutate(request_date = as_date(creation_timestamp)) %>% 
+  group_by(request_date, official_suburb) %>% 
+  count() %>% 
+  ungroup()
+
+tmp %>% 
+  ggplot(aes(request_date, n, fill = official_suburb))+
+  geom_col()+
+  facet_wrap(~official_suburb)+
+  scale_fill_manual(values = coct_cols)
+
+x <- tmp %>% 
+  arrange(desc(n)) %>% 
+  slice(1:4) %>% 
+  pull(request_date)
+
+```
+
+A closer look at the codes for those dates, reveals traffic congestion, and traffic light failures as the causes.
+
+```{r}
+
+top_ol <- c("BRACKENFELL COMMON", "MILNERTON",
+            "PARKLANDS")
 
 dta_dir %>% 
-  filter(official_suburb %in% x) %>% 
+  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
+  filter(creation_timestamp %in% x &
+           official_suburb %in% top_ol) %>% 
+  group_by(creation_timestamp, official_suburb, code) %>% 
+  count() %>% 
+  ungroup() %>% 
+  arrange(desc(n)) %>% 
+  filter(n != 1) %>% 
+  kable(format = "simple") %>% 
+  kable_styling(position = "left")
+
+```
+
+If those four dates are excluded in a sensitivity analysis, the top requesting suburbs table ranking changes(Table x).
+
+```{r}
+tmp <- dta_dir %>% 
+  mutate(across(creation_timestamp, ~as_date(.x))) %>% 
+  filter(!creation_timestamp %in% x)
+
+x <- tmp %>% 
+  group_by(directorate) %>% 
+  count() %>%
+  pull(n)
+
+tmp %>% 
+  group_by(official_suburb) %>% 
+  count() %>% 
+  ungroup() %>% 
+  mutate(percentage = round(n/x*100,2)) %>% 
+  arrange(desc(percentage)) %>% 
+  slice(1:5) %>% 
+  kable(format = "simple")
+
+```
+
+Taken together, the results highlight Marconi Beam, Table View and Milnerton, as contributors of the highest number of requests for the majority of the months included here.
+A breakdown of the codes associated with requests for the top 3 suburbs is given in Fig.
+x.
+
+```{r}
+# explore codes for top 
+select_suburbs <- c("MARCONI BEAM", "TABLE VIEW", "MILNERTON")
+
+dta_dir %>% 
+  filter(official_suburb %in% select_suburbs) %>% 
   group_by(official_suburb, code) %>% 
@@
   ungroup() %>% 
-  ggplot(aes(n, code, fill = official_suburb))+
-  geom_col(position = "dodge")
-
-# top 3
-
-select_suburbs <- c("PARKLANDS", "MILNERTON")
-
-```
-
-## 1. Focusing on the Urban Mobility directorate - "What is the median & 80th percentile time to complete each service request across the City?" (each row represent a service request).
+  ggplot(aes(n, reorder(code,n), fill = official_suburb))+
+  geom_col(position = "dodge")+
+  scale_fill_manual(values = coct_cols)+
+  labs(y = "Code")+
+  theme(axis.title.y = element_blank())
+
+```
+
+It is important to note that the top 3 suburbs were identified from the available data.
+The current dataset does not provide the opportunity to compare the tendency of the citizens in a particular suburb to report faults or make requests.
+Higher numbers of requests in a suburb, may be the sum of a greater tendency to report, and an average amount of infrastructure challenges, so that suburbs with the highest number of actual infrastructure problems, may be underrepresented in the data, due to under-reporting.
+
+## Service request completion times
 
@@
 
+```
+
+The median and 80th percentile of request completion times for each suburb, as well as for all suburbs combined, were calculated.
+From the city-wide numbers it can be seen that most requests were completed within hours, while it took almost a week to complete 80% of some requests.
+
+```{r}
+
 # look at suburbs with n >= 5
@@
            n >= n_cutoff) %>% 
-  kable(format = "simple", caption = "City-wide, all codes combined") %>%
+  kable(format = "simple", caption = "City-wide, all codes combined", digits = 2) %>%
   kable_styling( position = "left")
 
+```
+
+Because it is clear that the completion time data is skewed, with most tasks taking hours to days, and a minority, taking weeks, it is useful to look at requests taking less than, or more than 24 hours to complete, separately.
+
+```{r}
+x <- time_dta_list$all_times %>% 
+  filter(official_suburb == "ALL COMBINED" &
+           code == "All codes combined") %>% 
+  mutate(perc = round(x/n*100,2)) %>% 
+  pull(perc)
+
+```
+
+Requests completed in less than 24 hours, represents `r x`% of all requests.
+
+```{r}
 # filter by < 24 hrs
@@
            n >= n_cutoff) %>% 
-  kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in less than 24 hours") %>%
+  kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in less than 24 hours",
+        digits = 2) %>%
   kable_styling( position = "left")
@@
            n >= n_cutoff) %>% 
-  kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in 24 hours or more") %>%
+  mutate(across(c(med_time_hours, q80_time_hours),
+                ~.x/24)) %>% 
+  rename(med_time_days = med_time_hours, 
+         q80_time_days = q80_time_hours) %>% 
+  kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in 24 hours or more",
+        digits = 2) %>%
   kable_styling( position = "left")
 
+```
+
+Suburbs with the top 5 longest median completion times are shown, as well as suburbs with the top 5 shortest median completion times.
+
+```{r}
 # split by 24 hours
-
-x <- time_dta_list$below_24 %>% 
-  filter(official_suburb == "ALL COMBINED" &
-           code == "All codes combined") %>% 
-  pull(n)
-
-x <- time_dta_list$all_times %>% 
-  filter(official_suburb == "ALL COMBINED" &
-           code == "All codes combined") %>% 
-  mutate(perc = x/n*100) %>% 
-  pull(perc)
-
-# x
 
@@
   arrange(desc(med_time_hours)) %>% 
+  mutate(across(c(med_time_hours, q80_time_hours),
+                ~.x/24)) %>% 
+  rename(med_time_days = med_time_hours, 
+         q80_time_days = q80_time_hours) %>% 
   slice(1:5) %>% 
-  kable(format = "simple", caption = "Top 5 suburbs with longest median completion times, all codes combined") %>%
+  kable(format = "simple", caption = "Top 5 suburbs with longest median completion times, all codes combined",
+        digits = 2) %>%
   kable_styling( position = "left")
@@
 
+```
+
+Looking at combined completion times for all codes, probably masks important information, as some due to the nature of some codes, they are more likely to take longer than others.
+Exploring any codes for which there were at least 5 service requests, reveals that different types of codes differ in their typical duration times (Fig.
+x).
+The x-axis of the plot is on a log base 10 scale, to aid in visualisation.
+
+```{r}
+
+sub_txt <- "ALL COMBINED"
+code_txt <- "All codes combined"
+n_cut_off <- 5
+
+time_dta_list$all_times %>% 
+  filter(n >= n_cutoff & official_suburb == sub_txt &
+           code != code_txt) %>% 
+  pivot_longer(c(med_time_hours, q80_time_hours),
+               names_to = "quantile", 
+               values_to = "hours") %>% 
+  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
+   geom_line(colour = "grey70")+
+  geom_point(size = 3)+
+   scale_colour_manual(values = coct_cols)+
+  theme(axis.title.y = element_blank())
+
+```
+
+Breaking down the codes into the suburbs they originated from, and including only codes with 5 or more requests in a suburb, displays the variation present in the completion times, across suburbs.
+Fig.
+x is split into completion times below or above 24 hours.
+The variation is probably enhanced due to the relatively small size of the dataset, and exploring these times over more months, will give greater insight.
+
+```{r}
 # split by codes
@@
                   heights = c(2,1))
-
-tmp <- map(x,
-           ~time_dta_list$all_times %>% 
-  filter(official_suburb != sub_txt &
-           code == .x &
+# 
+# tmp <- map(x,
+#            ~time_dta_list$all_times %>% 
+#   filter(official_suburb != sub_txt &
+#            code == .x &
+#            n >= n_cutoff) %>% 
+#   arrange(desc(med_time_hours))) 
+# 
+# x <- seq(1,length(tmp), by = 2)
+# 
+# kable(tmp[1:2], format = "simple") %>%
+#     kable_styling( position = "left")
+# 
+# 
+```
+
+Focusing on the three suburbs identified before as focus-areas, it is shown that the median and 80th percentile for completion times, are shorter for all three suburb, compared to all suburbs combined.
+
+```{r}
+
+sub_txt <- select_suburbs
+code_txt <- "All codes combined"
+n_cut_off <- 5
+
+time_dta_list$all_times %>% 
+  filter(official_suburb %in% "ALL COMBINED" &
+           code == code_txt) %>% 
+  bind_rows(time_dta_list$all_times %>% 
+  filter(official_suburb %in% sub_txt &
+           code == code_txt)) %>% 
+  arrange(desc(med_time_hours)) %>% 
+  kable(format = "simple")
+
+```
+
+Plotting the median and 80th percentile completion times for each code that represented 5 or more requests, shows that a limited set of codes make up the majority of requests for these suburbs, and that completion time varies according to code.
+
+```{r}
+time_dta_list$all_times %>% 
+  filter(official_suburb %in% sub_txt &
+           code != code_txt &
            n >= n_cutoff) %>% 
-  arrange(desc(med_time_hours)))
-
-x <- seq(1,length(tmp), by = 2)
-
-kable(tmp[1:2], format = "simple") %>%
-    kable_styling( position = "left")
-
-# per month
-
-# weekends, public holidays?
-
-```
-
-## 2.
-Focusing on the Urban Mobility directorate - "What is the median & 80th percentile time to complete each service request for the 3 suburbs identified in (1)?" (each row represent a service request).
+  pivot_longer(c(med_time_hours, q80_time_hours),
+               names_to = "quantile", 
+               values_to = "hours") %>% 
+  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
+   geom_line(colour = "grey70")+
+  geom_point(size = 3)+
+  facet_wrap(~official_suburb)+
+   scale_colour_manual(values = coct_cols)+
+  theme(axis.title.y = element_blank())
+
+```
 
@@
 
-```{r}
-
-time_dta_list$all_times %>% 
-  filter(official_suburb %in% c(select_suburbs, "ALL COMBINED"))
-
-# wilcox test
-map(select_suburbs,
-    ~dta_dir %>% 
-  select(official_suburb, time_hours) %>%
-  mutate(suburb_bin = case_when(official_suburb %in% 
-                                  .x ~ .x,
-                                TRUE ~ "ALL COMBINED")) %>% 
-  summarise(suburb = .x,
-            p_value = wilcox.test(time_hours ~ 
-                                    suburb_bin)$p.value))
-
-# prop.test
+To determine whether the differences in median and 80th percentile completion times for the selected suburbs are different from the completion times of all other suburbs combined, each completion time of a suburb, was rated as either lower or, equal to or higher, than the median completion time for all other suburbs.
+The proportion of completion times lower than the median of all others combined, was then compared to the proportion of the completion times of all other suburbs combined, which was lower than the median (as this would work out to roughly 0.5, by definition, the test is essentially a comparison of the suburb of interest proportion, to 0.5).
+The proportions were compared with a statistical test for equality of proportions, and a p-value < 0.05, denotes that the probability that a suburb would have a proportion so different from all other suburbs, by random chance alone (i.e.
+no underlying reason for the difference), is less than 5%.
+The test was significant for bith Marconi Beam and Milnerton, but not for Table View (Table x).
+
+```{r}
+
+# time_dta_list$all_times %>% 
+#   filter(official_suburb %in% c(select_suburbs, "ALL COMBINED")) 
+# 
+# # wilcox test
+# map(select_suburbs,
+#     ~dta_dir %>% 
+#   select(official_suburb, time_hours) %>%
+#   mutate(suburb_bin = case_when(official_suburb %in% 
+#                                   .x ~ .x,
+#                                 TRUE ~ "ALL COMBINED")) %>% 
+#   summarise(suburb = .x,
+#             p_value = wilcox.test(time_hours ~ 
+#                                     suburb_bin)$p.value))
+# 
+# # prop.test
 
@@
 
-prop_test_list
+prop_test_list %>% 
+  bind_rows() %>% 
+  arrange(p_value) %>% 
+  kable(format = "simple")
 
@@
 ## 4. Identify value-adding insights for the management of Urban Mobility, from the dataset provided, in regard to commuter transport within the City.
+
+## Other relevant insights
 

### 2023-08-13 draft_v0_1.Rmd
Dashboard mock-up added. From Windows.

@@
 
+# fix typo in code
+
+sort(unique(dta$code[dta$directorate == "URBAN MOBILITY"]))
+dta$code <- gsub("Conjested", "Congested", dta$code) 
 ```
@@
 
-## 4. Identify value-adding insights for the management of Urban Mobility, from the dataset provided, in regard to commuter transport within the City.
+```{r}
+
+theme_tmp <- theme_doc+
+  theme(text = element_text(size = 9))
+theme_set(theme_tmp)
+# top suburbs
+
+x <- nrow(dta_dir)
+
+tmp <- dta_dir %>% 
+  group_by(official_suburb) %>% 
+  count() %>% 
+  ungroup() %>% 
+  mutate(percentage = (n/x*100)) %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5)
+
+p1 <- tmp %>% 
+  ggplot(aes(percentage, reorder(official_suburb, percentage),
+             fill = official_suburb))+
+  geom_col()+
+  scale_fill_manual(values = coct_cols)+
+  labs(title = "Top 5 Suburbs Contributing Requests")+
+  theme(legend.title = element_blank(),
+        axis.title.y = element_blank())
+
+top5 <- tmp$official_suburb
+
+p2 <- dta_dir %>% 
+  group_by(month_creation) %>% 
+  count(name = "total_month") %>% 
+  inner_join(dta_dir %>% 
+  group_by(month_creation, official_suburb) %>% 
+  count()) %>% 
+  ungroup() %>% 
+  filter(official_suburb %in% top5) %>% 
+  mutate(percentage = (n/total_month*100)) %>% 
+  arrange(desc(percentage)) %>% 
+  ggplot(aes(month_creation, percentage, colour = official_suburb,
+             group = official_suburb))+
+  geom_point(size = 3)+
+  geom_line(linewidth = 2)+
+  theme(legend.title = element_blank(),
+        axis.title.y = element_blank())+
+  scale_colour_manual(values = coct_cols)+
+  labs(title = "Top 5 Suburbs Contributing Requests, Per Month")
+
+# top codes for top 5 suburbs
+
+n_cutoff <- 5
+
+tmp <- dta_dir %>% 
+  filter(official_suburb %in% top5) %>% 
+  group_by(month_creation, official_suburb) %>% 
+  count(name = "total_month") %>% 
+  ungroup() %>% 
+  inner_join(dta_dir %>% 
+  filter(official_suburb %in% top5) %>% 
+  group_by(month_creation, official_suburb, code) %>% 
+  count() %>% 
+  ungroup() %>% 
+  filter(n >= n_cutoff)) %>% 
+  mutate(percentage = n/total_month*100)
+
+p3 <- tmp %>% 
+  ggplot(aes(percentage, reorder(code, percentage),
+             fill = month_creation))+
+  geom_col(position = "dodge")+
+  facet_wrap(~official_suburb, scales = "free_y",
+             ncol = 1)+
+  scale_fill_manual(values = coct_cols)+
+  theme(legend.title = element_blank(),
+        axis.title.y = element_blank())
+
+A <- ggpubr::ggarrange(p1, p2, common.legend = TRUE, ncol =1)
+A <- ggpubr::ggarrange(A, p3, nrow = 1)
+
+# completion times
+
+  # flag things from past three months with missing completion times
+  # create dataset as if it is 1 April 2020
+
+dta_apr <- dta_dir %>% 
+  mutate(date_complete = as_date(completion_timestamp),
+         date_create = as_date(creation_timestamp),
+        across(c(completion_timestamp, date_complete, month_completion),~ifelse(month_completion %in% c("jan", "feb", "mar"),
+                                .x, NA)),
+        date_today = as_date("2020-04-01"))
+
+tmp <- dta_apr  %>% 
+  filter(is.na(completion_timestamp)) %>% 
+  mutate(interv = date_create %--% date_today,
+          time_elapsed_weeks = as.duration(interv) / dweeks(1)) %>% 
+  select(-interv) %>% 
+  filter(time_elapsed_weeks > 4)
+
+p1 <- tmp %>% 
+  group_by(code) %>% 
+  count() %>% 
+  ungroup() %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5) %>% 
+  ggplot(aes(n, reorder(code, n)))+
+  geom_col(fill = coct_cols[2])+
+  theme(axis.title.y  = element_blank())+
+  labs(title = "Top 5 Codes Pending Requests for 4 Weeks or More")
+
+p2 <- tmp %>% 
+  group_by(official_suburb) %>% 
+  count() %>% 
+  ungroup() %>% 
+  arrange(desc(n)) %>% 
+  slice(1:5) %>% 
+  ggplot(aes(n, reorder(official_suburb, n)))+
+  geom_col(fill = coct_cols[1])+
+  theme(axis.title.y  = element_blank())+
+  labs(title = "Top 5 Suburbs Pending Requests for 4 Weeks or More")
+
+B <- ggpubr::ggarrange(p1, p2, ncol = 1)
+
+# city-wide completion time med, q80
+
+sub_txt <- "ALL COMBINED"
+code_txt <- "All codes combined"
+n_cut_off <- 5
+
+p1 <- time_dta_list$all_times %>% 
+  filter(n >= n_cutoff & official_suburb == sub_txt &
+           code != code_txt) %>% 
+  pivot_longer(c(med_time_hours, q80_time_hours),
+               names_to = "quantile", 
+               values_to = "hours") %>% 
+  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
+   geom_line(colour = "grey70")+
+  geom_point(size = 3)+
+   scale_colour_manual(values = coct_cols)+
+  theme(axis.title.y = element_blank())
+
+p2 <- time_dta_list$all_times %>% 
+  filter(n >= n_cutoff & official_suburb != sub_txt &
+           code != code_txt) %>% 
+  arrange(desc(med_time_hours)) %>% 
+  slice(1:5) %>% 
+  pivot_longer(c(med_time_hours, q80_time_hours),
+               names_to = "quantile", 
+               values_to = "hours") %>% 
+  ggplot(aes(log10(hours), reorder(code,hours), group = code, colour = quantile))+
+   geom_line(colour = "grey70")+
+  geom_point(size = 3)+
+   scale_colour_manual(values = coct_cols)+
+  theme(axis.title.y = element_blank())+
+  facet_wrap(~reorder(official_suburb,hours), scale = "free_y", ncol = 1)
+
+C <- ggpubr::ggarrange(p1, p2, nrow = 1)
+
+D <- ggpubr::ggarrange(A,  B, labels = c("A", "B"))
+E <- ggpubr::ggarrange(C, labels = "C")
+
+ggpubr::ggarrange(D, E, ncol = 1)
+
+theme_set(theme_doc)
+
+knit_exit()
+
+```
 
@@
 
+Traffic congestion can be explored by plotting request numbers associated with this code, over time, and looking for patterns in day of week, or highlighting particularly bad days (Fig.
+x).
+Several days in February 2020 stand out.
+
 ```{r}
@@
 
- unique(dta_dir$code)
-
- code_txt <- "Conjested Traffic at Intersection"
+ code_txt <- "Congested Traffic at Intersection"
 
@@
 x <- sort(tmp$n, decreasing = TRUE)
-x <- x[6]
+x <- x[6] # for top 5
 
@@
                            paste(day_of_week, creation_timestamp),
-                         TRUE ~ "")) %>%
+                         TRUE ~ ""),
+       weekend = case_when(day_of_week %in% c("Sat", "Sun") ~ "weekend",
+                           TRUE ~ "week")) %>%
 
  ggplot(aes(creation_timestamp, n, group = code,
-            label = label))+
+            label = label, colour = weekend))+
  geom_line(colour = "grey70")  +
- geom_point()+
-   geom_text()
-
+ geom_point(size = 3)+
+   geom_text()+
+  scale_color_manual(values = coct_cols)+
+  theme(legend.title = element_blank())
+
+```
+
+The suburbs contributing to the top 5 days with highest number of traffic congestion requests, show that a small number of suburbs make the largest contribution.
+
+```{r fig.height = 16}
 dta_dir %>% 
@@
   ggplot(aes(n, reorder(official_suburb,n), fill = as.factor(create_date)))+
-  geom_col()
+  geom_col()+
+  scale_fill_manual(values = coct_cols)+
+  labs(fill = "Top 5 Dates")+
+  theme(axis.title.y = element_blank())
+
+# get outlier date
+tmp <- dta_dir %>% 
+  mutate(date = as_date(creation_timestamp)) %>% 
+  filter(code == code_txt & official_suburb == "PARKLANDS") %>% 
+  group_by(date) %>% 
+  count() %>% 
+  ungroup() %>% 
+  arrange(desc(n)) %>% 
+  slice(1) %>% 
+  pull(date)
+
+```
+
+A clear pattern for number of requests and day of the week, is seen for congested traffic requests, and to some degree for Traffic lights out.
+Wednesdays contributed the most traffic congestion requests, but when removing the outlier for Parklands on `r tmp`,there seems to be a mid-week plateau, tapering down to Sunday and Saturday on either side, as expected.
+
+```{r}
 
@@
 
- dta_dir %>% 
-   filter(code %in% x) %>% 
+p1 <- dta_dir %>% 
+  mutate(date = as_date(creation_timestamp)) %>% 
+   filter(code %in% x ) %>% 
  mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
@@
    ggplot(aes(day_of_week, n, group = code, colour = code))+
- geom_point()+
- geom_line()
-
- dta_dir %>% 
+ geom_point(size = 3)+
+ geom_line()+
+   scale_colour_manual(values = coct_cols)+
+  labs(title = "All values")
+
+p2 <- dta_dir %>% 
+  mutate(date = as_date(creation_timestamp)) %>% 
+  filter(!(date == tmp & code == code_txt)) %>% 
+   filter(code %in% x ) %>% 
  mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
- filter(code == code_txt) %>% 
- group_by(day_of_week) %>% 
- count() %>% 
- ggplot(aes(day_of_week, n))+
- geom_col()
-
- # ?stats
- # odds of having congestion on a wednesday vs other weekdays
-
+ group_by(code, day_of_week) %>% 
+ count() %>%
+   ggplot(aes(day_of_week, n, group = code, colour = code))+
+ geom_point(size = 3)+
+ geom_line()+
+   scale_colour_manual(values = coct_cols)+
+  labs(title = "Parklands outlier removed")
+
+ggpubr::ggarrange(p1, p2, common.legend = TRUE)
+
+ # 
+ # dta_dir %>% 
+ # mutate(day_of_week = wday(creation_timestamp, label=TRUE)) %>% 
+ # filter(code == code_txt) %>% 
+ # group_by(day_of_week) %>% 
+ # count() %>% 
+ # ggplot(aes(day_of_week, n))+
+ # geom_col()
+
+```
+
+The pattern is less pronounced when looking at individual suburbs, but the three suburbs identified previously for intervention, feature prominently in the highest numbers of requests(Fig x).
+
+```{r}
  # suburbs and days of week
@@
    ggplot(aes(day_of_week, n, label = label))+
-   geom_point(alpha = 0.6)+
-   geom_text(alpha = 0.5, size = sz, check_overlap = TRUE)
-
-```
+   geom_jitter(alpha = 0.6, width = 0.1)+
+   ggrepel::geom_text_repel(alpha = 0.5, size = sz,
+                            max.overlaps = 50,
+                            force = 0.1)
+
+```
+
+## Conclusion
 

### 2023-08-14 draft_v0_1.Rmd
final draft. From Windows.

@@
 **Note from the author:  This report was compiled to complete the City of Cape Town (COCT) data science challenge.  It is not an official COCT report.**  
-The goal of this analysis is to highlight COCT suburbs that the Urban Mobility Directorate should focus on, for infrastructure improvements, based on service request data for 4 months (December 2019 to March 2020).
+The goal of this analysis is to highlight COCT suburbs that the Urban Mobility Directorate should focus on, for infrastructure improvements, based on service request data for 3 months (January 2020 to March 2020).
 It also explored intervals between creation and completion of service requests, to assess performance across the city.
@@
 The ```sr_hex_truncated.csv``` dataset was provided for this analysis.
-It originates from service requests across COCT spanning December 2019 to March 2020.
+It originates from service requests across COCT spanning 31 December 2019 to March 2020.
 The analysis was performed in the Statistical Programming Language R, using a RMarkdown file to produce this document.
-The code is available in a Github repository () and this analysis should be fully reproducible.
+The code is available in a Github repository (https://github.com/EliznaM/ds_code_challenge.git) and this analysis should be fully reproducible.
 The R and package versions used are listed at the end of this document in an appendix.
@@
 Medians, quantiles and non-parametric tests for differences in service request completion times were used because of skew data distribution.
-Only 1 day of 2019 is represented in the data (31 December) and it was excluded from the analysis.
+For the same reason, log base 10 transformation was sometimes used to aid visualisations.
+Only 1 day of 2019 is represented in the data (31 December), therefore it was excluded from the analysis.
 Missing data entries were explored and details focusing on missingness can be found in the appendix at the end of this document.
@@
   mutate(percentage_missing = n_missing/total_ym*100) %>% 
-  kable(format = "simple") %>%
-   kable_styling(position = "left")
+  kable(caption = "", format = "simple")
 
@@
 appx_list$miss_cause <- tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of {select_directorate} entries.")) %>%
-  kable(col.names = NULL) %>%
-  kable_styling(position = "left")
+  kable(caption = "", format = "simple")
 
@@
   filter(missing_completion_time != 0) %>% 
-  kable(format = "simple") %>%
-  kable_styling(position = "left")
+  kable(caption = "", format = "simple")
 
@@
 This section provides broad context of service requests across all COCT Directorates for the time period included in the dataset.
-The table displays number of service requests per Directorate, as well as the percentage each represents of the total requests for all Directorates.
+Table 1 displays number of service requests per Directorate, as well as the percentage each represents of the total requests for all Directorates.
 The Urban Mobility Directorate requests make up a small percentage of total requests.
@@
 tmp %>% 
-  kable(format = "simple", caption = "Number and percentage of entries per Directorate") %>%
-  kable_styling( position = "left") %>% 
-  row_spec(row = x, bold = TRUE)
-
-```
-
-The percentages of total requests seem to remain fairly stable per month, in the included time period, with Water and Sanitation and Energy contributing the bulk of requests.
+  kable(format = "simple", caption = "Number and percentage of entries per Directorate")
+
+```
+
+The percentages of total requests seem to remain fairly stable per month, in the included time period, with Water and Sanitation and Energy contributing the bulk of requests (Fig. 1).
 
@@
   facet_wrap(~year_creation, scales = "free_x")+
-  labs(fill = "Directorate, ordered from smallest to largest contribution",
+  labs(title = "Figure 1",
+    fill = "Directorate, ordered from smallest to largest contribution",
        x = "Month", y = "Percentage of total requests")
@@
 
-Highlighting the top 3 suburbs (largest percentage of total per suburb) for each Directorate, reveals that for most Directorates the total requests are made up of many suburbs, contributing relatively low proportions to the total.
+Highlighting the top 3 suburbs (largest percentage of total per suburb) for each Directorate, reveals that for most Directorates the total requests are made up of many suburbs, contributing relatively low proportions to the total (Fig.
+2).
 For Urban Mobility, for example, the biggest contribution is Milnerton, around 4%, which means that many other suburbs contribute percentages lower than 4%, to make up the total of requests for this Directorate.
@@
 
-```{r fig.height=12, fig.width=24}
+```{r fig.height=8, fig.width=24}
 map(unique(tmp$directorate),
@@
   scale_fill_manual(values = c(coct_cols, coct_cols))+
-  labs(x = "Percentage")+
+  labs(title = "Fig. 2",
+    x = "Percentage")+
   theme(legend.position = "none",
@@
 This section provides insight to inform decisions for infrastructure improvement efforts.
-It is important to note that the suburb information was missing for a large percentage of service requests (Table x).
+It is important to note that the suburb information was missing for a large percentage of service requests (Table 2).
 
@@
   mutate(percentage_missing_suburb = round(total_missing/total_ym*100, 2)) %>% 
-  kable(format = "simple") %>%
-  kable_styling( position = "left")
+  kable(caption = "Missing suburb entries", format = "simple")
 
@@
 The reason for the suburb missingness is not clear from the data.
-For the top 5 codes with most service requests, "Pothole&Defect Road Foot Bic Way/Kerbs" and "Paint Markings Lines&Signs" were more likely to be missing, than not (Table x).
+For the top 5 codes with most service requests, "Pothole&Defect Road Foot Bic Way/Kerbs" and "Paint Markings Lines&Signs" were more likely to be missing, than not (Table 3).
 
@@
   rename(not_missing = `FALSE`, missing = `TRUE`) %>% 
-  kable(format = "simple")
-
-```
-
-The odds of a suburb entry being missing was more than double for a number of dates, the top 4 of which occur closely together in March 2020 (Table x).
+  kable(caption = "Missing suburbs, by code", format = "simple")
+
+```
+
+The odds of a suburb entry being missing was more than double for a number of dates, the top 4 of which occur closely together in March 2020 (Table 4).
 There may therefore be a systematic reason for this missingness, but it is not clear from this dataset.
@@
   slice(1:5) %>% 
-  kable(format = "simple") %>%
-  kable_styling( position = "left")
+  kable(caption = "Missing suburbs, by date", format = "simple")
 
@@
   slice(1:5) %>% 
-  summarise(perc = sum(n)/x*100)
+  summarise(perc = round(sum(n)/x*100, 2))
 
@@
 An important action point may therefore be to attempt to find the cause for the large proportion of missing suburb entries.
-To identify the top suburbs to focus infrastructure improvements on, one can take different approaches.  Simply taking the total number of service requests, would highlight the suburbs in Table x.
+To identify the top suburbs to focus infrastructure improvements on, one can take different approaches.  Simply taking the total number of service requests, would highlight the suburbs in Table 5.
 
@@
 tmp %>% 
-  kable(format = "simple") %>% 
-  kable_styling(position = "left")
+  kable(caption = "Total requests", format = "simple")
 
@@
 One can also look at consistent numbers of requests over the 3 month period(Fig.
-x).
+3).
 Of the top 5 by simple totals, Milnerton, Marconi Beam and Table View are in the top 5 per month, three, two and two times, respectively, while Parklands and Brackenfell Common appear only once.
 
-```{r}
+```{r fig.height=10}
 # nr over time, per suburb, missing suburbs excluded now
@@
         axis.title.y = element_blank())+
-  scale_fill_manual(values = coct_cols)
-
-```
-
-Calculating the percentage of requests each suburb represents, per month, and taking the median of those percentages over the three included months, produces Table x.
+  scale_fill_manual(values = coct_cols)+
+  labs(title = "Fig. 3")
+
+```
+
+Calculating the percentage of requests each suburb represents, per month, and taking the median of those percentages over the three included months, produces Table 6.
 
@@
 tmp %>% 
-  kable(format = "simple") %>%
-  kable_styling( position = "left")
+  kable(caption = "Totals by suburb", format = "simple")
 
@@
 It can also be seen that Milnerton was responsible for an unusually high percentage in February, and Marconi Beam in March.
-Exploring request numbers at a more granular level, reveals that 4 particular dates have a disproportionate influence on the high request numbers in Brackenfell Common, Milnerton and Parklands.
+Exploring request numbers at a more granular level, reveals that 4 particular dates have a disproportionate influence on the high request numbers in Brackenfell Common, Milnerton and Parklands (Fig.
+4).
 
@@
   facet_wrap(~official_suburb)+
-  scale_fill_manual(values = coct_cols)
+  scale_fill_manual(values = coct_cols)+
+  labs(title = "Figure 4",
+       caption = "Number of service requests is on y-axis and dates 
+       on the x-axis, while each panel represents a different suburb.")
 
@@
 
-A closer look at the codes for those dates, reveals traffic congestion, and traffic light failures as the causes.
+A closer look at the codes for those dates, reveals traffic congestion, and traffic light failures as the causes (Table 7).
 
@@
   filter(n != 1) %>% 
-  kable(format = "simple") %>% 
-  kable_styling(position = "left")
-
-```
-
-If those four dates are excluded in a sensitivity analysis, the top requesting suburbs table ranking changes(Table x).
+  kable(caption = "Outier dates", format = "simple")
+
+```
+
+If those four dates are excluded in a sensitivity analysis, the top requesting suburbs table ranking changes(Table 8).
 
@@
   slice(1:5) %>% 
-  kable(format = "simple")
+  kable(caption = "Suburb totals, outliers excluded", format = "simple")
 
@@
 A breakdown of the codes associated with requests for the top 3 suburbs is given in Fig.
-x.
+5.
 
@@
   scale_fill_manual(values = coct_cols)+
-  labs(y = "Code")+
+  labs(title = "Figure 5", y = "Code")+
   theme(axis.title.y = element_blank())
@@
 The current dataset does not provide the opportunity to compare the tendency of the citizens in a particular suburb to report faults or make requests.
-Higher numbers of requests in a suburb, may be the sum of a greater tendency to report, and an average amount of infrastructure challenges, so that suburbs with the highest number of actual infrastructure problems, may be underrepresented in the data, due to under-reporting.
-
-## Service request completion times
+Higher numbers of requests in a suburb, may be the sum of a greater tendency to report, and the amount of infrastructure challenges, so that suburbs with the highest number of actual infrastructure problems, may be underrepresented in the data, due to under-reporting.
+
+## Service Request Completion Times
 
@@
 
-The median and 80th percentile of request completion times for each suburb, as well as for all suburbs combined, were calculated.
+The median and 80th percentile of request completion times for each suburb, as well as for all suburbs combined, were calculated (Table 9, med = median; q80 = 80th percentile).
 From the city-wide numbers it can be seen that most requests were completed within hours, while it took almost a week to complete 80% of some requests.
@@
            n >= n_cutoff) %>% 
-  kable(format = "simple", caption = "City-wide, all codes combined", digits = 2) %>%
-  kable_styling( position = "left")
-
-```
-
-Because it is clear that the completion time data is skewed, with most tasks taking hours to days, and a minority, taking weeks, it is useful to look at requests taking less than, or more than 24 hours to complete, separately.
+  kable(format = "simple", caption = "City-wide, all codes combined", digits = 2)
+
+```
+
+Because it is clear that the completion time data is skewed, with most tasks taking hours to days, and a minority, taking weeks, it is useful to look at requests taking less than, or more than 24 hours to complete, separately (Table 10 and 11, med = median; q80 = 80th percentile).
 
@@
   kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in less than 24 hours",
-        digits = 2) %>%
-  kable_styling( position = "left")
+        digits = 2)
 
@@
   kable(format = "simple", caption = "City-wide, all codes combined, only requests completed in 24 hours or more",
-        digits = 2) %>%
-  kable_styling( position = "left")
-
-```
-
-Suburbs with the top 5 longest median completion times are shown, as well as suburbs with the top 5 shortest median completion times.
+        digits = 2)
+
+```
+
+Suburbs with the top 5 longest median completion times are shown, as well as suburbs with the top 5 shortest median completion times (Table 12 and 13, med = median; q80 = 80th percentile).
 
@@
   kable(format = "simple", caption = "Top 5 suburbs with longest median completion times, all codes combined",
-        digits = 2) %>%
-  kable_styling( position = "left")
+        digits = 2)
 
@@
   slice(1:5) %>% 
-  kable(format = "simple", caption = "Top 5 suburbs with shortest median completion times, all codes combined") %>%
-  kable_styling(position = "left")
-
-```
-
-Looking at combined completion times for all codes, probably masks important information, as some due to the nature of some codes, they are more likely to take longer than others.
+  kable(format = "simple", caption = "Top 5 suburbs with shortest median completion times, all codes combined") 
+```
+
+Looking at combined completion times for all codes, probably masks important information, as due to the nature of some codes, they are more likely to take longer than others.
 Exploring any codes for which there were at least 5 service requests, reveals that different types of codes differ in their typical duration times (Fig.
-x).
+6).
 The x-axis of the plot is on a log base 10 scale, to aid in visualisation.
@@
    scale_colour_manual(values = coct_cols)+
-  theme(axis.title.y = element_blank())
+  theme(axis.title.y = element_blank())+
+  labs(title = "Figure 6",
+       caption = "X-axis shows log base 10 transformed hours,
+       so that 1 denotes 10 hours, 2, 100 hours, etc. Median
+       time is shown in pink and 80% percentile in green.")
 
@@
 Fig.
-x is split into completion times below or above 24 hours.
+7 is split into completion times below or above 24 hours.
+The coloured bars denote quantiles in the data, so that the innermost yellow section contains 50% of the data, the next colour stretches to contain 80% data, and the outer edges represent 95% of the data.
 The variation is probably enhanced due to the relatively small size of the dataset, and exploring these times over more months, will give greater insight.
@@
     ggdist::stat_interval()+
-  facet_wrap(~code, scales = "free_y", ncol = 1)
+  facet_wrap(~code, scales = "free_y", ncol = 1)+
+  labs(title = "Figure 7")
 
@@
 
-Focusing on the three suburbs identified before as focus-areas, it is shown that the median and 80th percentile for completion times, are shorter for all three suburb, compared to all suburbs combined.
+Focusing on the three suburbs identified before as focus-areas, it is shown that the median and 80th percentile for completion times, are shorter for all three suburbs, compared to all suburbs combined (Table 14, med = median, q80 = 80th percentile).
 
@@
   arrange(desc(med_time_hours)) %>% 
-  kable(format = "simple")
-
-```
-
-Plotting the median and 80th percentile completion times for each code that represented 5 or more requests, shows that a limited set of codes make up the majority of requests for these suburbs, and that completion time varies according to code.
-
-```{r}
+  kable(caption = "Request completion times", format = "simple")
+
+```
+
+Plotting the median and 80th percentile completion times for each code that represented 5 or more requests, shows that a limited set of codes make up the majority of requests for these suburbs, and that completion time varies according to code (Fig.
+8).
+
+```{r fig.height=10}
 time_dta_list$all_times %>% 
@@
    scale_colour_manual(values = coct_cols)+
-  theme(axis.title.y = element_blank())
-
-```
-
-## 3.
-"Is there any significant differences in the median and 80th percentile completion times between the City as a whole and the 3 suburbs identified in(1)?".
-Please elaborate on the similarities or differences.
+  theme(axis.title.y = element_blank())+
+  labs(title = "Figure 8", caption = "X-axis shows log base 10 transformed hours,
+       so that 1 denotes 10 hours, 2, 100 hours, etc. Median
+       time is shown in pink and 80% percentile in green.")
+
+```
 
 To determine whether the differences in median and 80th percentile completion times for the selected suburbs are different from the completion times of all other suburbs combined, each completion time of a suburb, was rated as either lower or, equal to or higher, than the median completion time for all other suburbs.
-The proportion of completion times lower than the median of all others combined, was then compared to the proportion of the completion times of all other suburbs combined, which was lower than the median (as this would work out to roughly 0.5, by definition, the test is essentially a comparison of the suburb of interest proportion, to 0.5).
+The proportion of completion times lower than the overall median, was then compared to the proportion of the completion times of all other suburbs combined, which was lower than the median (as this would work out to roughly 0.5, by definition, the test is essentially a comparison of the suburb of interest proportion, to 0.5).
 The proportions were compared with a statistical test for equality of proportions, and a p-value < 0.05, denotes that the probability that a suburb would have a proportion so different from all other suburbs, by random chance alone (i.e.
 no underlying reason for the difference), is less than 5%.
-The test was significant for bith Marconi Beam and Milnerton, but not for Table View (Table x).
+The test was significant for both Marconi Beam and Milnerton, but not for Table View (Table 15).
 
@@
   arrange(p_value) %>% 
-  kable(format = "simple")
-
-```
-
-## 3.
-Provide a visual mock of a dashboard for the purpose of monitoring progress in applying the insights developed in (1) & (2).
-It should focus the user on performance pain points.
-Add a note for each visual element, explaining how it helps fulfill this overall function.
-Please also provide a brief explanation as to how the data provided would be used to realise what is contained in your mock.
+  kable(caption = "Difference test p-values", format = "simple")
+
+```
+
+## Dashboard Mock-up
+
+A dashboard can help to identify performance pain points.
+Below is a rough mock-up of plots that could be displayed to give a single-page overview.
+The mock-up is based on a scenario, where data for the preceding 3 completed months are available to display in the dashboard.
+Each section (A to C) is described.
+Section A highlights suburbs with high numbers of service requests.
+Totals over three months are shown in a barplot, and change over time in a line plot.
+Furthermore the top 5 contributing codes to these numbers, are shown for every suburb, with contributing percentages.
+The data used for this section will need to be prepared carefully to identify exceptionally high  numbers (outliers) contributed by only a few isolated dates.
+These dates can then be removed, or imputed with the median of all other dates, to prevent such outliers from an exaggerated influence on the overall message.
+Section B focuses on pending service requests from the preceding three months.
+The barplots are produced by filtering for all service requests with missing completion times, and calculating the time from request creation, to the date on which the dashboard is accessed, giving the pending time.
+The codes and suburbs contributing to pending requests are shown separately.
+Section C reveals service request completion times.
+Missing completion times will be filtered out, therefore it is important to represent these in Section B, otherwise Section C will hide problems with pending tasks.
+As it has been shown that different codes and suburbs vary in completion times, a breakdown by code and suburb is given, highlighting the top 5 suburbs with longest median completion times.
+As missing data can influence interpretation, variables or suburbs with large proportions of missing data can potentially also be flagged on this dashboard, so that data quality can continuously be improved.
 
@@
 
-knit_exit()
-
-```
-
-## Other relevant insights
+```
+
+## Other Relevant Insights
 
 Traffic congestion can be explored by plotting request numbers associated with this code, over time, and looking for patterns in day of week, or highlighting particularly bad days (Fig.
-x).
+9).
 Several days in February 2020 stand out.
@@
   scale_color_manual(values = coct_cols)+
-  theme(legend.title = element_blank())
-
-```
-
-The suburbs contributing to the top 5 days with highest number of traffic congestion requests, show that a small number of suburbs make the largest contribution.
+  theme(legend.title = element_blank())+
+  labs(title = "Figure 9",
+       caption = "X-axis shows dates. Dots are coloured by week day (pink) or weekend day (green).
+       Outlier points are labelled with dates.")
+
+```
+
+The suburbs contributing to the top 5 days with highest number of traffic congestion requests, show that a small number of suburbs make the largest contribution (Fig. 10).
 
@@
   labs(fill = "Top 5 Dates")+
-  theme(axis.title.y = element_blank())
+  theme(axis.title.y = element_blank())+
+  labs(title = "Figure 10",
+       caption = "Each bar show number of requests, with different colours for each date's contribution.")
 
@@
 A clear pattern for number of requests and day of the week, is seen for congested traffic requests, and to some degree for Traffic lights out.
-Wednesdays contributed the most traffic congestion requests, but when removing the outlier for Parklands on `r tmp`,there seems to be a mid-week plateau, tapering down to Sunday and Saturday on either side, as expected.
+Wednesdays contributed the most traffic congestion requests, but when removing the outlier for Parklands on `r tmp`,there seems to be a mid-week plateau, tapering down to Sunday and Saturday on either side, as expected (Fig.
+11).
 
@@
    scale_colour_manual(values = coct_cols)+
-  labs(title = "All values")
+  labs(title = "Figure 11, All values")
 
@@
    scale_colour_manual(values = coct_cols)+
-  labs(title = "Parklands outlier removed")
+  labs(title = "Parklands outlier removed",
+       caption = "Each coloured line is a separate code.")
 
@@
 
-The pattern is less pronounced when looking at individual suburbs, but the three suburbs identified previously for intervention, feature prominently in the highest numbers of requests(Fig x).
+The pattern is less pronounced when looking at individual suburbs, but the three suburbs identified previously for intervention, feature prominently in the highest numbers of requests(Fig 12).
 
@@
    ggplot(aes(day_of_week, n, label = label))+
-   geom_jitter(alpha = 0.6, width = 0.1)+
+   geom_jitter(alpha = 0.6, width = 0.1, colour = coct_cols[1])+
    ggrepel::geom_text_repel(alpha = 0.5, size = sz,
                             max.overlaps = 50,
-                            force = 0.1)
+                            force = 0.1)+
+   labs(title = "Figure 12",
+        caption = "Each dot represents a suburb, only ones with
+        high numbers are labelled.")
 
@@
 ## Conclusion
+
+Three suburbs for the Urban Mobility Directorate to potentially focus infrastructure improvement efforts on, were highlighted, although it must be noted that the current data does not allow for exploration of under-reporting in some suburbs.
+Service requests were typically completed within hours, but some codes were completed in longer times, so that it took almost a week to complete 80% of requests.
+Exploring traffic congestion related requests, revealed that the middle 3 days of the week are worse than Mondays and Fridays, and suburbs on the West Coast are particularly challenged.
 
@@
 # save in appendix list
-appx_list <- list()
 
@@
 
-appx_list$miss_dir <- dta %>% 
-  group_by(official_suburb) %>% 
-  count(name = "total_suburb") %>% 
-  inner_join(dta %>% 
-  filter(is.na(directorate) & !is.na(official_suburb)) %>% 
-  group_by(official_suburb) %>% 
-  count(name = "n_missing")) %>% 
-  mutate(percentage_missing = n_missing/total_suburb*100) %>% 
-  filter(n_missing >= n_cutoff) %>% 
-  arrange(desc(percentage_missing), desc(n_missing)) %>% 
-  ungroup()
-
-# max(tmp$percentage_missing)
-# tmp$official_suburb[tmp$percentage_missing == max(tmp$percentage_missing)]
-
-appx_list$upset_data <- dta %>% 
-  select(directorate, department, branch, section, official_suburb) %>% 
-  mutate(across(everything(), ~as.numeric(is.na(.x)))) %>% 
-  as.data.frame()
-
-appx_list$mis_dir_nonmis_sub <- appx_list$upset_data %>% 
-  group_by(directorate, official_suburb) %>% 
-  count() %>% 
-  filter(directorate == 1 & official_suburb == 0) %>% 
-  pull(n)
+tmp1 <- appx_list$miss_dir 
+mis_dir_nonmis_sub <- appx_list$mis_dir_nonmis_sub
 
@@
 
-appx_list$na_interpret_affected <- UpSetR::upset(tmp)
-
-na_interpret_affected
+appx_list$na_interpret_affected
 grid::grid.text("Fig. A",x = 0.65, y=0.95, gp=grid::gpar(fontsize=20))
@@
 n_cutoff <- 5
-plotdta <- dta %>% 
-  group_by(official_suburb) %>% 
-  count(name = "total_suburb") %>% 
-  inner_join(dta %>% 
-  filter(is.na(directorate) & !is.na(official_suburb)) %>% 
-  group_by(official_suburb) %>% 
-  count(name = "n_missing")) %>% 
-  mutate(percentage_missing = n_missing/total_suburb*100) %>% 
-  filter(n_missing >= n_cutoff) %>% 
-  arrange(desc(percentage_missing)) %>% 
-  ungroup()
-
-appx_list$plot_mis_dir_nonmis_sub <- plotdta %>% 
-  slice(1:20) %>% 
-  ggplot()+
-  geom_point(aes(percentage_missing, reorder(official_suburb, percentage_missing), size = n_missing))+
-  labs(title = "Fig. B: Top 20 Suburbs with missing Directorate")
-
+
+appx_list$plot_mis_dir_nonmis_sub 
 # missing directorate and dates
 
-appx_list$table_mis_dir_dates <- dta %>% 
-  group_by(year_creation, month_creation) %>% 
-  count(name = "total_ym") %>% 
-  inner_join(dta %>% 
-               group_by(year_creation, month_creation, is.na(directorate)) %>% 
-               count(name = "n_missing")) %>%
-  ungroup() %>% 
-  rename(missing_directorate = 4) %>% 
-  filter(missing_directorate == TRUE) %>% 
-  mutate(percentage_missing = n_missing/total_ym*100) %>% 
-  kable(format = "simple") %>%
-   kable_styling(position = "left")
-
-```
-
-Of the entries with Official suburb missing, ...
-
-```{r}
-
-plotdta <- dta %>% 
-  group_by(directorate) %>% 
-  count(name = "total_directorate") %>%
-  inner_join(dta %>% 
-               filter(is.na(official_suburb)) %>% 
-               group_by(directorate) %>% 
-                        count(name = "n_missing")) %>% 
-  ungroup() %>% 
-  mutate(percentage_missing = n_missing/total_directorate*100)
-
-appx_list$mis_sub <- plotdta %>% 
-  ggplot()+
-  geom_point(aes(percentage_missing, reorder(directorate, percentage_missing), size = n_missing))+
-  labs(title = "Fig. C: Missing Suburbs per Directorate")
-
+appx_list$table_mis_dir_dates
+
+```
+
+Most Directorates had relatively large percentages of entries with Official Suburb missing.
+
+```{r}
+
+appx_list$mis_sub 
 ```
@@
 
-dta_dir <- dta %>% 
-  filter(directorate == select_directorate)
-
-x <- which(names(dta_dir) %in% c("code_group", "code", "cause_code_group", "cause_code"))
-
-# map(dta_dir[x], ~sort(unique(.x))) 
-# map(dta_dir[x], ~sum(is.na(.x))) 
-# 
-# dta_dir %>%
-#   select(all_of(x)) %>% 
-#   group_by(is.na(.)) %>% 
-#   count()
-
-x <- nrow(dta_dir)
-
-x1 <- dta_dir %>% 
-  group_by(mis_code_grp = is.na(cause_code_group), 
-           mis_cause_code = is.na(cause_code)) %>% 
-  count(name = "missing_cause") %>% 
-  ungroup() %>% 
-  filter(mis_code_grp == TRUE & mis_cause_code == TRUE) %>% 
-  pull(missing_cause)
-
-appx_list$miss_cause <- tibble(msg = glue::glue("Missing cause for {round(x1/x*100, 2)}% of {select_directorate} entries.")) %>%
-  kable(col.names = NULL) %>%
-  kable_styling(position = "left")
+appx_list$miss_cause
 
@@
 
-appx_list$miss_compl_dates <- dta_dir %>% 
-  group_by(official_suburb) %>% 
-  summarise(n = n(),
-            missing_completion_time = 
-              sum(is.na(completion_timestamp))) %>% 
-  filter(missing_completion_time != 0) %>% 
-  kable(format = "simple") %>%
-  kable_styling(position = "left")
-
-```
+appx_list$miss_compl_dates
+
+```
+
+Session info
 

### 2023-08-14 draft_v0_1.Rmd
Final draft typos and text sizes fixed. Dataset outside folder deleted. From Linux.

@@
   theme(text = element_text(color = "grey30",
-                            size = 18),
+                            size = 19),
         line = element_line(color = "grey30"),
         rect = element_rect(color = "grey30",
-                            fill = "#FFFFF6"),
-        panel.background = element_rect(fill = "#FFFFF6", colour = "grey70"),
+                            fill = "#FFFFF2"),
+        panel.background = element_rect(fill = "#FFFFF2", colour = "grey70"),
         # axis.line = element_blank(),
@@
         strip.background = element_rect(colour = "grey70",
-                                        fill = "#FFFFF6"),
+                                        fill = "#FFFFF2"),
         strip.text = element_text(colour = "grey30"))
@@
 
-appx_list$na_interpret_affected <- UpSetR::upset(appx_list$upset_data)
+appx_list$na_interpret_affected <- UpSetR::upset(appx_list$upset_data, text.scale = 3)
 
@@
 
-```{r}
+```{r overview2}
 dta %>% 
@@
 
-```{r fig.height=8, fig.width=24}
+```{r overview3,fig.height=8, fig.width=24}
 map(unique(tmp$directorate),
@@
 
-```{r focused5}
+```{r focused3}
 
@@
 
-```{r fig.height=10}
+```{r focused5, fig.height=10}
 # nr over time, per suburb, missing suburbs excluded now
@@
 
-```{r}
+```{r focused6}
 tmp <- dta_dir %>% 
@@
 
-```{r}
+```{r focused7}
 
@@
 
-```{r}
+```{r focused8}
 
@@
   filter(n != 1) %>% 
-  kable(caption = "Outier dates", format = "simple")
+  kable(caption = "Outlier dates", format = "simple")
 
@@
 
-```{r}
+```{r focused9}
 tmp <- dta_dir %>% 
@@
 
-```{r}
+```{r focused10}
 # explore codes for top 
@@
 
-```{r}
+```{r times1}
 
@@
 
-```{r}
+```{r times2}
 
@@
 
-```{r}
+```{r times3}
 
@@
 
-```{r}
+```{r times4}
 x <- time_dta_list$all_times %>% 
@@
 
-```{r}
+```{r times5}
 # filter by < 24 hrs
@@
 
-```{r}
+```{r times6}
 # split by 24 hours
@@
 
-```{r}
+```{r times7}
 
@@
 
-```{r}
+```{r times8}
 # split by codes
@@
 
-```{r}
+```{r times9}
 
@@
 
-```{r fig.height=10}
+```{r times10, fig.height=10}
 time_dta_list$all_times %>% 
@@
 
-```{r}
+```{r times11}
 
@@
 
-```{r}
+```{r dash}
 
@@
 
-```{r}
+```{r insights1}
 # traffic congestion over time
@@
 x <- x[6] # for top 5
-
+sz <- 5
 tmp %>% 
@@
  geom_point(size = 3)+
-   geom_text()+
+   geom_text(size = sz)+
   scale_color_manual(values = coct_cols)+
@@
 
-```{r fig.height = 16}
+```{r insights2,fig.height = 16}
 dta_dir %>% 
@@
 
-```{r}
+```{r insights3}
 
@@
 
-```{r}
+```{r insights4}
  # suburbs and days of week
@@
 
-sz <- 3
+sz <- 5
  dta_dir %>% 
@@
 
-```{r}
+```{r appx1}
 
@@
 
-```{r}
+```{r appx2}
 
@@
 
-```{r}
+```{r appx3}
 
@@
 
-```{r}
+```{r appx4}
 
@@
 
-```{r}
+```{r appx5}
 
@@
 
-```{r}
+```{r appx6}
 

### 2023-08-14 main_analysis.Rmd
Final analysis, main_analysis created. Draft versions archived.

--- 2023-08-14 draft_v0_1.Rmd
@@
 x <- x[6] # for top 5
-sz <- 5
+sz <- 7
 tmp %>% 
@@
 
-sz <- 5
+sz <- 7
  dta_dir %>% 
